---
title: "Appendix"
output: html_notebook
bibliography: cadre-appendix-library.bib
---

## Introduction

This is an early draft of what will become the companion Appendix to the modeling 
papers that we publish. Narrative text and the R code are interwoven. The computer code is also available in a public [GitHub repository](https://github.com/khanna7/cadre). 

This is written in  [R Markdown](http://rmarkdown.rstudio.com) to facilitate [reproducibility](https://bookdown.org/m_p_sperrin/intro_to_r_2020/projects-and-reproducibility.html), i.e. the document can be automatically updated whenever underlying code or data changes.
This is my first attempt at a full-scale reproducible document; 
please bear with me as I figure out kinks and assess if this would be a workable 
format for writing technical Appendices. 

```{r, results='hide', echo=FALSE, message=FALSE}
# Libraries ----------

library(ergm)
library(network)
library(dplyr)

```
 
## Population Overview

A synthetic population consisting of `r n` agents between 18 and 64 years of age is simulated in an agent-based model (ABM) with social network structure. This simulated population is broadly representative of the 18-64 year old population of RI. The population size is deliberately kept small in the early stages to make the model more  efficient in terms of computation time. 

Each individual is assigned a set of demographic characteristics (age, race, sex), behavioral profiles (smoking, alcohol use), incarceration history. Feedback loops between the impacts of incarceration on smoking and alcohol use are simulated in the criminal-justice involved persons, and their social networks.   

```{r, results='hide', echo=FALSE, message=FALSE}
# Initialize Population --------------

n <- 5000
net <- network.initialize(n, directed = FALSE)
```

## Demographic Characteristics

### Age
Per [census data](https://censusreporter.org/profiles/04000US44-rhode-island/), 
63% of the RI population is in the range of our interest (18-64 years). We consider the population to be uniformly distributed across this range, which might not be unreasonable given the shape of the distribution above.

```{r, results='hide', echo=TRUE, message=FALSE, fig.show='hide'}
age.min <- 18
age.max <- 64
age <- runif(n, age.min, age.max)
net %v% "age" <- age
hist(net %v% "age")
```


### Race
Per [census data](see https://www.census.gov/quickfacts/RI), we parameterize the population as belonging to 5 race/ethnic categories: White, Black, Hispanic, Asian, and Other.

```{r, results='hide', echo=TRUE, message=FALSE}
white_alone <- 71.4/100 #not Hispanic or Latino
black_alone <- 8.5/100
hispanic_alone <- 16.3/100
asian_alone <- 3.7/100
other <- 1- sum(c(white_alone, black_alone, asian_alone, hispanic_alone))

race_cats <- c("White", "Black", "Hispanic", "Asian", "Other")

race <- 
  sample(x=race_cats, 
       prob = c(white_alone, black_alone, hispanic_alone, asian_alone, other),
       replace = TRUE,
       size = n
       )
net %v% "race" <- race
table(net %v% "race")

```


### Sex

Per [census data](see https://www.census.gov/quickfacts/RI),   
the proportion of females in RI is `r female.prop*100`. 
We dichotomize `r female.prop*100`% of our our agents as being female, and
`r 100-female.prop*100`% as being male.


```{r, results='hide', echo=TRUE, message=FALSE}

# sex (https://www.census.gov/quickfacts/RI)
female.prop <- 51.3/100
female <- rbinom(n, 1, prob=female.prop)

net %v% "female" <- female
table(net %v% "female")
```


## Substance Use Behaviors

### Smoking

Per the [CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm), the national prevalence of current smokers in the US is `r smoking.prev*100`%. 
( The CDC defines "a current smoker" as someone who "who reported smoking at least 100 cigarettes during their lifetime and who, at the time they participated in a survey about this topic, reported smoking every day or some days.")

```{r, results='show', echo=TRUE, message=FALSE}
smoking.prev <- 0.13
net %v% "smoking.prob" <- smoking.prev
smoking.prob <- net %v% "smoking.prob"
```

The [CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm)
also stratifies the national smoking rates by age, sex, and race/ethnicity. 
There is also a geographic prevalence of smoking (about 10% in the northeastern US),
that  is not stratified by demographic variables.  



### Alcohol Use

We classify alcohol use behaviors in one of four categories, defined as in a previous pair of 
by frequency of use, as per @Apostolopoulos2018I; @Apostolopoulos2018II.
The four categories are: "abstainers", "occassional users", "regular users", "persons with high-risk use or Alcohol Use Disorder (AUD)". 

To parameterize the distribution of persons in each of these categories, 
data from the National Comorbidity Survey Replication, as published in 
@Kalaydjian2009, where:

    * (A) 91.7% (SE=0.9) reported >1 sip of alcohol at some time in their life, 
    * (B) 72.9% (1.3) reported using alcohol regularly at some time in their life
    * (C) 13.2% (0.6) met criteria for alcohol abuse at some time in their life, and 
    * (D) 5.4% (0.3) met criteria for alcohol dependence at some time in their life.
 
 Thus, the proportion of abstainers is $1-A = 8.3%$. For simplicity, 
 Category B above is taken to be "occasional users", Category C as "regular users". 
 Category D as High-Risk/AUD = 100-(A+B+C) = 5.6%, which is almost the same as 
 D reported by @Kalaydjian2009.

## Initiatializing the Incarceration Process in the Model

```{r, results='hide', echo=FALSE, message=FALSE}
# 2-week risk #macmadu 2021 (table 1)
incarceration.2week.prob <- 1.7/100
  # we'll break down the above by age and race
```

Incarceration fuctions by defining two locations: 
a "correctional setting" (CS), and a "home" setting. Per @Macmadu2021, the 
2-week incarceration probability for RI persons is 
`r incarceration.2week.prob*100`%. We thus initialize the model with `r incarceration.2week.prob*100`% of persons in the correctional setting.  

```{r, results='hide', echo=FALSE, message=FALSE}
# attributes:
  # current location (Correctional Setting vs Home)
  locations <- c("CS", "H")

  curr_loc <- 
    sample(x=locations, 
           prob = c(incarceration.2week.prob, 1-incarceration.2week.prob),
           replace = TRUE,
           size = n
    )
  
  incarcerated <- recode(curr_loc, "H" = "0", "CS" = "1")
  incarcerated <- as.numeric(incarcerated)
  table(incarcerated)
  
  net %v% "curr_loc" <- curr_loc
  table(net %v% "curr_loc")
  
  net %v% "incarcerated" <- incarcerated
  table(net %v% "incarcerated")
  
  # ever incarcerated
  ever_inc <- ifelse(curr_loc=="CS", 1, 0) 
  xtabs(~curr_loc + ever_inc)
  net%v% "ever_inc" <- ever_inc
    
  # number of times incarcerated
  num_incs <- ifelse(ever_inc == 1, 1, 0)
  xtabs(~num_incs + ever_inc)
  net %v% "num_incs" <- num_incs
  
  # other predictors of incarceration?
  
  # other parameters:
  #- sentence duration: see macmadu 2021 (Table 1) for male/female distribution
  # recidivism probability: as above
```
  

### Modeling Impact of Incarceration on Smoking


```{r, results='hide', echo=FALSE, message=FALSE}
# Black men with incarceration history 1.77x as likely to smoke as those w/o inc history
  # Black women with incarceration history 1.61x as likely 
  # (reference: Bailey 2015, AJPH)
  xtabs(~net %v% "race" + net %v%"smoker" +
          net %v% "female")

  mult.black.inc.male.smk <- 1.77 #see above
  mult.black.inc.female.smk <- 1.61 #see above
  
  black.male.idx <- intersect(which(net %v% "race" == "Black"), 
                                   which(net %v% "female" == 0)) 
  black.male.inc.idx <- intersect(black.male.idx, 
                                 (which(net %v% "ever_inc" == 1))) 
  
  black.female.idx <- intersect(which(net %v% "race" == "Black"), 
                             which(net %v% "female" == 1)) 
  black.female.inc.idx <- intersect(black.female.idx, 
                                 (which(net %v% "ever_inc" == 1))) 
  
  smoking.prob[black.male.inc.idx] <- smoking.prev * mult.black.inc.male.smk
  smoking.prob[black.female.inc.idx] <- smoking.prev * mult.black.inc.female.smk
  table(smoking.prob)
  net %v% "smoking.prob" <- smoking.prob
  
  # assign smoker status
  smoker <- rbinom(n, 1, smoking.prob)
  net %v% "smoker" <- smoker
  table(net %v% "smoker") #1= current smoker, 0=former/never smoker
  ```

Incarceration-induced stressors can increase the smoking by justice-involved persons. From data collected in  
the National Survey of Americal Life, @Bailey2015 report that Black men with incarceration history  `r{mult.black.inc.male.smk}` times as likely to smoke as those without incarceration
history, and Black women with incarceration history `r{mult.black.inc.female.smk} times as likely to smoke as women withoout incarceration history. For agents belonging to all race/ethnicities, 
we model increased odds of smoking as per the magnitudes reported by @Bailey2015.
(__Note: We may need to find data that stratify impact of incarceration on smoking
  by race/ethnicity__).


### Modeling Feedback between Incarceration and Alcohol Use

### Social Networks


## Next Steps

- Stratify current smokers by age, race, and sex

- Incarceration: Sentence duration and recidivism probability


### References
