---
title: "ALcohol and Incarceration Coding Problem 5/12"
author: "Yurui Zhang"
date: '2022-05-19'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Consider a population of 1000 agents. And the their behavior states of alcohol use can be described into four levels "Abstainers", "Occasional Drinkers", "Regular Drinkers", "High Risk/ AUD".

Based on our data, we know that 8.3% of our population is an abstainer, 72.9% of our population is an occasional drinker, 13.2/100% of our population is a regular drinker, and the rest are those in the high-risk/AUD (AUD is alcohol use disorder) category.

```{r}
library("markovchain")
```

###	Initialize a set of 1000 agents.

```{r}
x_start<-rep(c("Abstainers", "Occasional Drinkers", "Regular Drinkers", "High Risk/ AUD"),times=c(8.3/100*1000,72.9/100*1000,13.2/100*1000,5.6/100*1000))

set.seed(872436)           # Set seed
x_start <- sample(x_start)        # Sample vector

track_df <- data.frame(
   initial_start = x_start
)




```




```{r}

library(dplyr)

change <- function(times, start_state) {
   for(i in 1:times){
     for (k in 1:length(start_state)){
       if (nth(start_state, k) == "Abstainers"){
         if (runif(1,0,1)>0.9){
           start_state<-replace(start_state, k, "Occasional Drinkers")
         }
       }
       else if (nth(start_state, k) == "Occasional Drinkers"){
         if (runif(1,0,1)<0.05){
           start_state<-replace(start_state, k, "Abstainers")
         }
         else if (runif(1,0,1)>0.85){
           start_state<-replace(start_state, k,  "Regular Drinkers")
         }
       }
       else if (nth(start_state, k) == "Regular Drinkers"){
         if (runif(1,0,1)<0.05){
           start_state<-replace(start_state, k,  "Occasional Drinkers")
         }
         else if (runif(1,0,1)>0.85){
           start_state<-replace(start_state, k,  "High Risk/ AUD")
         }
       }
       else if (nth(start_state, k) == "High Risk/ AUD"){
         if (runif(1,0,1)<0.05){
           start_state<-replace(start_state, k,  "Regular Drinkers")
         }
       }
     }
   }
  return(start_state)
}
  



m <- change(10000, x_start) 

track_df$change_10000 <- m
track_df
```


###	Classify each agent initially as belonging to one of these four categories. 



###	Model a system where any person has a 10% chance at any time of “levelling up” in their alcohol use (e.g., an abstainer can become an occasional drinker, or a regular drinker can become “high risk/AUD”).

```{r}

alcoholStates <- c("Abstainers", "Occasional Drinkers", "Regular Drinkers", "High Risk/ AUD")
byRow = TRUE
alcoholMatrix <- matrix(data = c(0.99, 0.01, 0, 0,
                                0.01, 0.98, 0.01, 0,
                                0, 0.01, 0.98, 0.01,
                                0, 0, 0.01, 0.99), byrow = byRow, nrow = 4,dimnames = list(alcoholStates, alcoholStates))


mcalcohol <- new("markovchain", states = alcoholStates, byrow = byRow,
transitionMatrix = alcoholMatrix, name = "alcohol")
mcalcohol
```

```{r}
initialState <- c(1,0,0,0)
after2Days <- initialState * (mcalcohol * mcalcohol)
after7Days <- initialState * (mcalcohol^7)
after100Days <- initialState * (mcalcohol^100)
after10000Days <- initialState * (mcalcohol^10000)

after2Days
after7Days
after100Days
after10000Days
```

8.3% of our population is an abstainer, 72.9% of our population is an occasional drinker, 13.2% of our population is a regular drinker, and the rest are those in the high-risk/AUD (AUD is alcohol use disorder) category. 5.6%

### Metropolis Hastings Stationary Distribution


```{r}
library(expm)
matrix_x <- matrix(c(27/166,1/2,0,28/83,83/1458,1243/1458,132/1458,0,0,1/2,38/132,28/132,1/2,0,1/2,0), nrow = 4)

matrix_x
matrix_tt <- matrix_x %^% 1000

matrix_tt
```

```{r}

alcoholStates <- c("Abstainers", "Occasional Drinkers", "Regular Drinkers", "High Risk/ AUD")
byRow = TRUE



# need to be discussed

alcoholMatrix <- matrix(data = c(27/166, 1/2, 0, 28/83,
                                83/1458, 1243/1458, 132/1458, 0,
                                0, 1/2, 38/132, 28/132,
                                1/2, 0, 1/2, 0), byrow = byRow, nrow = 4,dimnames = list(alcoholStates, alcoholStates))


mcalcohol <- new("markovchain", states = alcoholStates, byrow = byRow,
transitionMatrix = alcoholMatrix, name = "alcohol")
mcalcohol
```


```{r}
initialState <- c(1,0,0,0)
after2Days <- initialState * (mcalcohol * mcalcohol)
after7Days <- initialState * (mcalcohol^7)
after100Days <- initialState * (mcalcohol^100)
after10000Days <- initialState * (mcalcohol^10000)

after2Days
after7Days
after100Days
after10000Days
```



###	Model a system where any person has a 5% chance at any time of “levelling down” in their alcohol use. 





