---
title: "Smoke Level Transition Matrix Analysis"
author: "Yurui Zhang and Aditya Khanna"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, results='hide', echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      include = TRUE,
                      results='hide')
```

# White Men

```{r}
white_Male_current <- 0.119
white_Male_former <- 0.598

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (white_Male_former/white_Male_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="white male", xlab = "current_stays", ylab = "former_stays")


grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)


former_stays
current_stays

tail(cbind(former_stays, current_stays), 200)

```

A reasonable pair of probability values for white men to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
white_male_optim_pair_value <- 987
white_male_no_trans_prob <- cbind(former_stays[white_male_optim_pair_value], 
                                  current_stays[white_male_optim_pair_value]) 
names(white_male_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
white_male_no_trans_prob

```
# White Women

```{r}
rm(list=c("former_stays", "current_stays"))

white_Female_current <- 0.104
white_Female_former <- 0.340

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (white_Female_former/white_Female_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="white Female", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)

```

A reasonable pair of probability values for white women to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
white_female_optim_pair_value <- 985
white_female_no_trans_prob <- cbind(former_stays[white_female_optim_pair_value], current_stays[white_female_optim_pair_value]) 
names(white_female_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
white_female_no_trans_prob
```
# Black Men

```{r}
rm(list=c("former_stays", "current_stays"))
black_Male_current <- 0.205
black_Male_former <- 0.473

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (black_Male_former/black_Male_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="black male", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)

```
A reasonable pair of probability values for Black men to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
black_male_optim_pair_value <- 985
black_male_no_trans_prob <- cbind(former_stays[black_male_optim_pair_value], current_stays[black_male_optim_pair_value]) 
names(black_male_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
black_male_no_trans_prob
```
# Black Women

```{r}
rm(list=c("former_stays", "current_stays"))
black_Female_current <- 0.113
black_Female_former <- 0.290

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (black_Female_former/black_Female_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="black Female", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)
```
A reasonable pair of probability values for Black women to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
black_female_optim_pair_value <- 985
black_female_no_trans_prob <- cbind(former_stays[black_female_optim_pair_value], current_stays[black_female_optim_pair_value]) 
names(black_female_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
black_female_no_trans_prob
```
# Hispanic Men

```{r}
rm(list=c("former_stays", "current_stays"))
hispanic_Male_current <- 0.127
hispanic_Male_former <- 0.492

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (hispanic_Male_former/hispanic_Male_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="hispanic male", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)

```
A reasonable pair of probability values for hispanic men to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
hispanic_male_optim_pair_value <- 985
hispanic_male_no_trans_prob <- cbind(former_stays[hispanic_male_optim_pair_value], current_stays[hispanic_male_optim_pair_value]) 
names(hispanic_male_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
hispanic_male_no_trans_prob
```
# Hispanic Women

```{r}
rm(list=c("former_stays", "current_stays"))
hispanic_Female_current <- 0.066
hispanic_Female_former <- 0.221

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (hispanic_Female_former/hispanic_Female_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="hispanic Female", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)
```

A reasonable pair of probability values for hispanic women to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
hispanic_female_optim_pair_value <- 985
hispanic_female_no_trans_prob <- cbind(former_stays[hispanic_female_optim_pair_value], current_stays[hispanic_female_optim_pair_value]) 
names(hispanic_female_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
hispanic_female_no_trans_prob

```
# Asian Men

```{r}
rm(list=c("former_stays", "current_stays"))
asian_Male_current <- 0.069
asian_Male_former <- 0.462

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (asian_Male_former/asian_Male_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="asian male", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)
```

A reasonable pair of probability values for Asian men to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
asian_male_optim_pair_value <- 985
asian_male_no_trans_prob <- cbind(former_stays[asian_male_optim_pair_value], current_stays[asian_male_optim_pair_value]) 
names(asian_male_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
asian_male_no_trans_prob
```

# Asian Women

```{r}
rm(list=c("former_stays", "current_stays"))
asian_Female_current <- 0.038
asian_Female_former <- 0.169

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (asian_Female_former/asian_Female_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="asian Female", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)

```
A reasonable pair of probability values for Asian women to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
asian_female_optim_pair_value <- 986
asian_female_no_trans_prob <- cbind(former_stays[asian_female_optim_pair_value], current_stays[asian_female_optim_pair_value]) 
names(asian_female_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
asian_female_no_trans_prob
```
# Other Men

```{r}
rm(list=c("former_stays", "current_stays"))
other_Male_current <- 0.155
other_Male_former <- 0.536

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (other_Male_former/other_Male_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="other male", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)

```
A reasonable pair of probability values for Other men to stay
as former and current smokers is: 


```{r, include=TRUE, results='markup'}
other_male_optim_pair_value <- 986
other_male_no_trans_prob <- cbind(former_stays[other_male_optim_pair_value], current_stays[other_male_optim_pair_value]) 
names(other_male_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
other_male_no_trans_prob
```

# Other Women
```{r}
rm(list=c("former_stays", "current_stays"))
other_Female_current <- 0.149
other_Female_former <- 0.410

former_stays <- seq(from = 0, to = 1, by = 0.001)
current_stays <- (other_Female_former/other_Female_current)*(former_stays-1)+1

plot(current_stays, former_stays, xlim=range(0:1), ylim=range(0:1),type = "b", pch = 19, col = "blue", main="other Female", xlab = "current_stays", ylab = "former_stays")

grid(nx = NULL, ny = NULL,
     col = "grey", lwd = 1)

former_stays
current_stays

tail(cbind(former_stays, current_stays), 100)

```
A reasonable pair of probability values for Other women to stay
as former and current smokers is: 

```{r, include=TRUE, results='markup'}
other_female_optim_pair_value <- 986
other_female_no_trans_prob <- cbind(former_stays[other_female_optim_pair_value], current_stays[other_female_optim_pair_value]) 
names(other_female_no_trans_prob) <- c("stay_former_smoker", "stay_current_smoker")
other_female_no_trans_prob
```

# Consolidated Results



```{r, include=TRUE, results='markup'}
smoking_maintain_status_matrix <- 
  as.matrix(
    rbind(
      white_male_no_trans_prob,
      white_female_no_trans_prob,
      black_male_no_trans_prob,
      black_female_no_trans_prob,
      hispanic_male_no_trans_prob,
      hispanic_female_no_trans_prob,
      asian_male_no_trans_prob,
      asian_female_no_trans_prob
    )
  )
row.names(smoking_maintain_status_matrix) <-
  c(
    "White Males",
    "White Females",
    "Black Males",
    "Black Females",
    "Hispanic Males",
    "Hispanic Females",
    "Asian Males",
    "Asian Females"
  )

colnames(smoking_maintain_status_matrix) <- 
  c("Maintain Former Status (p)", 
    "Maintain Current Status(p)"
  )


smoking_transition_matrix <-
  1-smoking_maintain_status_matrix
colnames(smoking_transition_matrix) <- 
  c("Former -> Current (p)",
    "Current -> Former (p)")
```

Consolidated probabilities for all race/ethnicity and gender combinations to stay
as former and current smokers are:

```{r echo=TRUE, results='markup'}
smoking_maintain_status_matrix
```
Consolidated probabilities former smokers to relapse and for current smokers
to cease smoking are:

```{r echo=TRUE, results='markup'}
knitr::kable(smoking_transition_matrix)
```

The average times to relapse are therefore:

```{r echo=TRUE, results='markup'}
smoking_transition_matrix
dim(smoking_transition_matrix)
smoking_transition_matrix[,1]
relapse_rates <- smoking_transition_matrix[,1]
smoking_time_to_relapse <-
  lapply(smoking_transition_matrix[,1], function (x) summary(rgeom(10000, x)))
smoking_time_to_relapse
```

The average times to cessation are:

```{r echo=TRUE, results='markup'}
cessation_rates <- smoking_transition_matrix[,2]
smoking_time_to_cessation <-
  lapply(smoking_transition_matrix[,2], function (x) summary(rgeom(10000, x)))
smoking_time_to_cessation
```

This analysis assumes that all current smokers  (conditioned on race and sex) 
are equally likely to try quitting and 
all former smokers (conditioned on race and sex) are equally likely to relapse.
  
This is probably not very realistic - a better assumption might consider agent-level 
behaviors that suggest that 55.1% of adult smokers made a quit attempt in the last
year and 7.5% quit ([CDC Fact Sheet](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/cessation/smoking-cessation-fast-facts/index.html)). 

See data from the PATH Study on relapse rates among persons who quit smoking  
[Pierce et al](https://paperpile.com/app/p/62406b1b-5617-08b9-8adf-52ddbd38b07b)
and relapse as a function of duration [Herd et al](https://paperpile.com/app/p/df6149a3-c109-0021-827d-90e99e20a1ab).


