---
title: "R Notebook"
output: html_notebook
---

We confirm below various attributes of the agents that are logged.

We will do this in `data.table` for speed. 

```{r}
library(data.table)
```

Read the agent data into data-table:

```{r}
agent_dt <- fread("/Users/adityakhanna/Google Drive/My Drive/code/cadre/python/output/agent_log_1.csv")
```

# Age Assignment at Entry

We check below the ages of new agents, 
and if their initial age assignments 
and age increments are working correctly. 

How many agents are present at tick 991:

```{r}
agent_dt[tick == 991, .N] 
```
How many agents present at tick 991 entered after model initialization?

```{r}
agent_dt[tick == 991 & agent_id > 9999, .N] 
```

The mean age of agents at time by time of entry at tick 1:
```{r}
agent_dt[tick == 1, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]


```
At tick 491:

```{r}
agent_dt[tick == 491, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]


```

And at tick 991:

```{r}
agent_dt[tick == 991, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]
```

These data seem to match expectations of aging of newly entering agents.
See that at tick 1, only one agent had entered after model initialization.
and this agent was 18 years old.

At tick 491, this agent was 19 years old (491-365 years older than 18, rounded), 
and at tick 991, this agent was 21 years old (991-365 years older than 18, rounded).

So, the aging process seems to work as expected.


# Incarceration Statistics

Mean number of incarcerations at tick 991:

```{r}
agent_dt[tick == 991, 
         .(
           "mean_n_incarcerations" = mean(n_incarcerations),
           "min_n_incarcerations" = min(n_incarcerations),
           "max_n_incarcerations" = max(n_incarcerations),
           "median_n_incarcerations" = median(n_incarcerations),
           "n_ever_incarcerated_agents"=sum(n_incarcerations > 0),
         .N)] 
```   
# Incarceration Duration

Compute statistics on agents whose last release date is 
after their last incarceration date.

```{r}
agent_dt[tick == 991 & last_release_tick > last_incarceration_tick, 
         .(
           "mean_duration" = mean(last_release_tick - last_incarceration_tick),
           "max_duration" = max(last_release_tick - last_incarceration_tick),
           "min_duration" = min(last_release_tick - last_incarceration_tick),
           "median_duration" = median(last_release_tick - last_incarceration_tick),
         .N)] 
```



# Release Statistics

## Number

```{r}
agent_dt[tick == 991, 
         .(
           "mean_n_releases" = mean(n_releases),
           "min_n_releases" = min(n_releases),
           "max_n_releases" = max(n_releases),
           "median_n_releases" = median(n_releases),
           "n_ever_released_agents"=sum(n_releases > 0),
         .N)] 
```
## Duration
Compute statistics on agents whose last release date is 
\*before\* their last incarceration date.

```{r}
agent_dt[tick == 991 & last_release_tick < last_incarceration_tick, 
         .(
           "mean_rel_dur" = mean(abs(last_release_tick - last_incarceration_tick)),
           "max_rel_dur" = max(abs(last_release_tick - last_incarceration_tick)),
           "min_rel_dur" = min(abs(last_release_tick - last_incarceration_tick)),
           "median_rel_dur" = median(abs(last_release_tick - last_incarceration_tick)),
         .N)] 
```
