---
title: "R Notebook"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

We confirm below various attributes of the agents that are logged.

We will do this in `data.table` for speed. 

```{r}
rm(list=ls())
library(data.table)
library(ggplot2)
library(yaml)
```

Read input parameters:¬

```{r}
input_params <- read_yaml("~/code/cadre/python/myparams/model_params.yaml")
```


Read the agent data into data-table:

```{r}
agent_dt <- fread("~/code/cadre/python/output/agent_log.csv")
str(agent_dt)
last_tick <- max(agent_dt$tick)
```

How many unique agents are in the dataset?

```{r}
length(unique(agent_dt$id))
range(unique(agent_dt$id))
```

There are `r length(unique(agent_dt$id))` unique agents in the dataset. Their IDs 
range from `r min((unique(agent_dt$id)))` to `r max((unique(agent_dt$id)))`.

# Demographics

```{r}
agent_dt_by_race <- agent_dt[tick == last_tick, .N, by=race][,"per_cent":=round(N/sum(N)*100)][order(race)]
target_values <- data.frame(race = names(input_params$RACE_DISTRIBUTION), 
                            target_pct = unlist(input_params$RACE_DISTRIBUTION))
```

```{r}
# Calculate the percentages
agent_dt_by_race[, .(count = .N), by = race][, percentage := round(count/sum(count) * 100, 2)]

# Order the data by race
agent_dt_by_race <- agent_dt_by_race[order(race)]

# Create the bar chart
ggplot(agent_dt_by_race, aes(x = race, y = per_cent, fill = race)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("#7bccc4", "#43a2ca", "#0868ac", "#f0f9e8")) +
  labs(title = "Race Distribution",
       x = "Race",
       y = "Percentage",
       fill = "Race")+
  geom_segment(data = target_values, 
               aes(x = race, xend = race, y = 0, yend = target_pct), 
               color = "red", size = 1, linetype = "dashed") +
  geom_text(data = target_values, 
            aes(x = race, y = target_pct*97, label = paste0(round(target_pct*100, 1), "%")),
            vjust = -0.5, size = 4, color = "red")
```

```{r echo=TRUE}
agent_dt[tick==last_tick, .N, by=female][,"%":=N/sum(N)][order(female)]
female_target <- input_params$FEMALE_PROP

gender_pct <- agent_dt[tick == last_tick, .N, by = female][, "%":= N/sum(N)][order(female)]
```
```{r}
# calculate percentages
gender_pct <- agent_dt[tick == last_tick, .N, by = female][, "%":= N/sum(N)][order(female)]

# read in YAML file with target value
female_target <- input_params$FEMALE_PROP

# calculate target percentages for each gender
female_target_pct <- female_target
male_target_pct <- 1 - female_target

# calculate actual percentages for each gender
female_actual_pct <- gender_pct[female == 1, `%`]
male_actual_pct <- gender_pct[female == 0, `%`]


# create plot
ggplot(gender_pct, aes(x = ifelse(female == 0, "Male", "Female"), y = `%`, fill = ifelse(female == 0, "Male", "Female"))) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02")) +
  annotate("text", x = 1, y = female_actual_pct, 
           label = paste0("Female Target = ", round(female_actual_pct*100, 1), "%"), 
           color = "#1b9e77", size = 4, vjust = 0) +
  annotate("text", x = 2, y = male_actual_pct, 
           label = paste0("Male Target = ", round(male_actual_pct*100, 1), "%"), 
           color = "#d95f02", size = 4, vjust = 0) +
  labs(title = "Agent Gender Distribution",
       x = "",
       y = "Percentage",
       fill = "") +
  theme_minimal()




```



```{r}
agebreaks <- c(18, 25, 35, 45, 55, 65)
agelabels <- c("18-24", "25-34", "35-44", "45-54", "55-64")
```


```{r}
agent_dt[tick == last_tick, .N, ]
setDT(agent_dt)[ , age_groups := cut(age, 
                                breaks = agebreaks, 
                                include.lowest = TRUE,
                                right = FALSE, 
                                labels = agelabels)]
nrow(agent_dt[tick == last_tick])
```

```{r}
agent_dt[tick == last_tick, .N, by=c("age_groups", "race", "female")][order(age_groups, race, female)]
agent_dt[tick == last_tick, .N, by=c("race", "age_groups", "female")][order(race, age_groups)]
```
```{r}
agent_dt[tick==last_tick, .N, by=c("race", "female")][,"%":=N/sum(N)*100][order(race)]
```

```{r}
agent_dt[tick==last_tick, .N, by=c("age_groups")][,"%":=round(N/sum(N)*100)][order(age_groups)]
```
Median age at the start: 

```{r}
agent_dt[tick==1, median(age)]
```

Median age at the end:

```{r}
nrow(agent_dt[tick==last_tick])
agent_dt[tick==last_tick, median(age)]
```

```{r}

# Create age groups
agent_dt_by_age_group <- agent_dt[tick==last_tick, .N, by = .(age_group = cut(age, breaks = c(18, 25, 35, 45, 55, 65, 75, 80, 84)))]

# Create bar chart
ggplot(agent_dt_by_age_group, aes(x = age_group, y = N, fill = age_group)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00", "#ffff33", "#a65628", "#f781bf")) +
  labs(title = "Agent Age Distribution",
       x = "Age Group",
       y = "Count",
       fill = "Age Group") +
  theme_minimal()

```

# Baseline Smoking Behaviors


At the first tick:

```{r}
agent_dt[tick==1 & race == "Black" & female == 0, .N, by=c("smoking_status")][]

agent_dt[tick==1 & race == "Black" & female == 0, .N, by=c("smoking_status")][,"sim_prop":= round(N/sum(N), 3)][]


# targets
BLACK_MALE_CURRENT = input_params$SMOKING_PREV$BLACK_MALE_CURRENT
BLACK_MALE_FORMER = input_params$SMOKING_PREV$BLACK_MALE_FORMER
BLACK_MALE_NEVER = input_params$SMOKING_PREV$BLACK_MALE_NEVER

smkg_prev_black_men_target <- c(BLACK_MALE_CURRENT, BLACK_MALE_FORMER, BLACK_MALE_NEVER)

smkg_prev_black_men_target

## Add difference between the two target and the sim

```

At the last tick:
```{r}
smkg_blackmen_last_tick <- agent_dt[tick==last_tick & race == "Black" & female == 0, .N, by=c("smoking_status")][,"prop":=round(N/sum(N), 3)][]
smkg_blackmen_last_tick

smkg_prev_black_men_target

```

```{r}
# Count the number and percentage of each smoking status
smoking_dt <- agent_dt[tick == last_tick, .N, by = smoking_status][, `:=` (percentage = round(N/sum(N)*100, 2))]
# Create the bar chart
ggplot(smoking_dt, aes(x = smoking_status, y = percentage, fill = smoking_status)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("#fec44f", "#43a2ca", "#7bccc4")) +
  labs(title = "Agent Smoking Status Distribution",
       x = "Smoking Status",
       y = "Percentage",
       fill = "Smoking Status") +
  theme_minimal()

```

# Baseline Alcohol Use

```{r}
alc_use_last_tick <- 
  agent_dt[tick == last_tick, .N, by = c("alc_use_status")][, 
  "%" := round(N /sum(N), 3)][
  order(alc_use_status)][
  ]

# targets
alc_use_props <- input_params$ALC_USE_PROPS
alc_use_targets <- c(alc_use_props$A, alc_use_props$O, alc_use_props$R, alc_use_props$D)

cbind(alc_use_last_tick[["%"]], alc_use_targets)
```

```{r}
# Calculate the percentages for alcohol use status
alc_use_dt <- agent_dt[tick == last_tick, .N, by = alc_use_status][, `:=` (percentage = round(N/sum(N)*100, 2))]
# Order the data table by alcohol use status
setorder(alc_use_dt, alc_use_status)
# Rename the alcohol use status codes to descriptive labels
alc_use_dt[, alc_use_status := factor(alc_use_status, levels = c(0, 1, 2, 3), 
                                      labels = c("Lifetime Abstainer", 
                                                 "Ocassional", 
                                                 "Regular", 
                                                 "Heavy"))]

ggplot(alc_use_dt, aes(x = alc_use_status, y = percentage, fill = alc_use_status)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Agent Alcohol Use Status Distribution",
       x = "Use Frequency",
       y = "Percentage",
       fill = "Alcohol Use Status") +
  theme_minimal() +
  guides(fill = "none") +
  geom_text(aes(label = paste0(as.numeric(input_params$ALC_USE_PROPS)*100, "%")), 
            position = position_stack(vjust = 1.02))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 10))


```

# Smoking and Alcohol Co-Use

```{r}
# Create a contingency table of smoking and alcohol use status
smoking_alc_use_table <- dcast(agent_dt[tick == last_tick], smoking_status ~ alc_use_status, value.var = "id", fun.aggregate = length)
smoking_alc_use_table <- smoking_alc_use_table[, .(smoking_status, `Lifetime Abstainer` = `0`, Occasional = `1`, Regular = `2`, Heavy = `3`)]
smoking_alc_use_melt <- melt(smoking_alc_use_table, id.vars = "smoking_status",
                             variable.name = "alcohol_use")
```

```{r}
smkg_alc_use_xtab <- dcast(smoking_alc_use_melt, smoking_status ~ alcohol_use, value.var = "value")
smkg_alc_use_xtab

smoking_alc_col_props <- prop.table(as.matrix(smkg_alc_use_xtab[,-1]), margin = 1)
smoking_alc_row_props <- prop.table(as.matrix(smkg_alc_use_xtab[,-1]), margin = 2)

smoking_alc_col_props <- cbind(smkg_alc_use_xtab[,1], round(smoking_alc_col_props, 3))
smoking_alc_row_props <- cbind(smkg_alc_use_xtab[,1], round(smoking_alc_row_props, 3))

print(smoking_alc_col_props)
print(smoking_alc_row_props)

rowSums(smoking_alc_col_props[,-1])
colSums(smoking_alc_row_props[,-1])
```
 

```
# Create a stacked bar chart of smoking and alcohol use status
ggplot(smoking_alc_use_melt, aes(x = smoking_status, y = value, fill = alcohol_use)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Agent Smoking and Alcohol Use Status Distribution",
       x = "Smoking Status",
       y = "Count",
       fill = "Alcohol Use Status") +
  theme_minimal() +
  scale_fill_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  guides(fill = "none")

```
```{r}
# Create a contingency table of smoking and alcohol use status
smoking_alc_use_table <- dcast(agent_dt[tick == last_tick], smoking_status ~ alc_use_status, value.var = "id", fun.aggregate = length)
smoking_alc_use_table <- smoking_alc_use_table[, .(smoking_status, `Lifetime Abstainer` = `0`, Occasional = `1`, Regular = `2`, Heavy = `3`)]
smoking_alc_use_melt <- melt(smoking_alc_use_table, id.vars = "smoking_status")

# Create a stacked bar chart of smoking and alcohol use status
ggplot(smoking_alc_use_melt, aes(x = smoking_status, y = value, fill = variable)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Agent Smoking and Alcohol Use Status Distribution",
       x = "Smoking Status",
       y = "Count") +
  scale_fill_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c"), 
                    labels = c("Lifetime Abstainer", "Occasional", "Regular", "Heavy"),
                    name = "Alcohol Use Status") +
  theme_minimal()

```
# Age Assignment at Entry

We check below the ages of new agents, 
and if their initial age assignments 
and age increments are working correctly. 

How many agents are present at the last tick:

```{r}
agent_dt[tick == last_tick, .N] 
```   
How many agents present at tick 991 entered after model initialization?

```{r}
agent_dt[tick == 991 & id > 9999, .N] 
```

The mean age of agents at time by time of entry at tick 1:
```{r}
agent_dt[tick == 1, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]


```
At tick 491:

```{r}
agent_dt[tick == 491, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]


```

And at tick 991:

```{r}
agent_dt[tick == 991, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]
```

These data seem to match expectations of aging of newly entering agents.
See that at tick 1, only one agent had entered after model initialization.
and this agent was 18 years old.

At tick 491, this agent was 19 years old (491-365 years older than 18, rounded), 
and at tick 991, this agent was 21 years old (991-365 years older than 18, rounded).

So, the aging process seems to work as expected.


# Incarceration Statistics

Number and proportion of incarcerated persons at last tick:

```{r}
agent_dt[tick==last_tick, .N, by=c("current_incarceration_status")][,"prop":=round(N/sum(N)*100, 0)][]
```
Break down number and proportions of incarcerated persons at last tick by race, sex and age:

```{r}
agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("race")][,"prop":=round(N/sum(N)*100, 0)][]

agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("female")][,"prop":=round(N/sum(N)*100, 0)][]

agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("age_groups")][,"prop":=round(N/sum(N)*100, 0)][]

agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("race", "female", "age_groups")][order(race, female, age_groups)][,"prop":=round(N/sum(N)*100, 0)][]
```

How many persons have been ever incarcerated?

To answer this question, we can group the last incarceration times by agent ID.
If last incarceration time is -1, the person has never been incarcerated. 
If last incarceration time is > -1, the person has been incarcerated at least once. 

```{r}
# Group data by agent ID and compute the last incarceration time
last_incarceration_time <- agent_dt[, 
                                    .(last_incarceration_time = max(last_incarceration_tick)), 
                                    by = id]

# Count the number of agents who have been incarcerated at least once
n_ever_incarcerated <- sum(last_incarceration_time$last_incarceration_time != -1)

# Compute the proportion of agents who have been incarcerated at least once
prop_ever_incarcerated <- n_ever_incarcerated / nrow(last_incarceration_time)


# Agents who have been incarcerated at least once
ever_incarcerated_times <- last_incarceration_time[last_incarceration_time > 1]
dim(ever_incarcerated_times)

# Extract demographic information for ever incarcerated agents
ever_incarcerated_ids <-
  unique(agent_dt[id %in% ever_incarcerated_times$id, id])
ever_incarcerated_info <- agent_dt[id %in% ever_incarcerated_ids,
                                   .SD[.N, list(
                                     id,
                                     age = last(age),
                                     race = last(race),
                                     female = last(female)
                                   )],
                                   by = id]

# Compute distributions
## mean age
age_summary <- ever_incarcerated_info[, .(mean_age=mean(age),
                                       min_age=min(age),
                                       max_age=max(age)
                                       )
                                   ]
sex_dist <-
  ever_incarcerated_info[, .N, by = female][,  .(
    Sex = c("Female", "Male"),
    Count = c(N[female == 0], N[female == 1]),
    "Proportion of ever incarcerated" = round(N / sum(N), 2)
  )]
print(sex_dist)

race_dist <- ever_incarcerated_info[, .N, by = race][,
  .(Race = race, 
    Count = N,
    Proportion = round(N / sum(N), 2)
   )
  ]
race_dist[]

```

Thus, `r n_ever_incarcerated` persons have been incarcerated at least once. The
proportion of persons who have been ever incarcerated is `r prop_ever_incarcerated`.

Now let us compute the number of females and males whoa have ever been incarcerated,
and the proportions by race (% of White, Black, ...)

```{r}
first_tick_race_sex <-
  agent_dt[, .SD[1], by = id][, .(
    ID = id,
    Tick = tick,
    Female = female,
    Race = race
  )]

# Count number of females and males
sex_count <- first_tick_race_sex[, .N, by = Female][,
                                               .(Sex = c("Female", "Male"), 
                                                 Count = c(N[Female == 1], N[Female == 0]))                                               ]
print(sex_count)

# Count number of individuals in each race
race_count <- first_tick_race_sex[, .N, by = Race][,
                                                   .(Race, Count = N)]
print(race_count)
```

Thus, the distribution of ever incarcerated persons by sex is

```{r}
cbind(sex_count$Sex, round(as.numeric(sex_dist$Count/sex_count$Count), 3))
```

The distribution of ever incarcerated persons by race is:

```{r}
cbind(race_count$Race, round(as.numeric(race_dist$Count/race_count$Count), 3))
```

What is the duration of incarceration for persons who have been ever incarcerated? 
Restricting our data to only persons who have been released from their last
incarceration episode, we can estimate this as:

```{r}
# Filter the data to include only the observations for agents who have been ever incarcerated and released
ever_incarcerated_released <-
  agent_dt[last_release_tick > last_incarceration_tick,
           .SD[1, ], # ensure times for each incarceration, release episode appear only once
           by = .(id, last_incarceration_tick)]

# Compute the length of incarceration time for each ever incarcerated and released agent
ever_incarcerated_released[, incarceration_length := last_release_tick - last_incarceration_tick,
                           by = id]

# Compute the summary statistics of the lengths of incarceration time for the ever incarcerated and released agents
incarceration_summary <-
  ever_incarcerated_released[, .(
    min_incarceration_length = min(incarceration_length),
    median_incarceration_length = median(incarceration_length),
    max_incarceration_length = max(incarceration_length)
  ),
  by = id]

# Compute the distribution of incarceration times for all incarceration episodes followed by release
summary(ever_incarcerated_released$incarceration_length)
```

Assess the total number of incarceration-release events and the number of unique
agents who experienced them:

```{r}
length(ever_incarcerated_released$id)
length(unique(ever_incarcerated_released$id))
```

Distribution of incarceration events:
```{r}
hist(ever_incarcerated_released$incarceration_length, 
     breaks = seq(0, 3000, by = 30),
     xlab = "Length of incarceration (days)",
     ylab = "Number of individuals",
     main = "Distribution of incarceration lengths")
```

Break down number and proportions of persons ever incarcerated at last tick by race, sex and age:

```{r}
ever_inc_dt <- 
  agent_dt[tick==last_incarceration_tick & n_incarcerations > 0, 
           .N, 
           by=c("race", "female", "age_groups")][order(race, female, age_groups)][
             ,"prop":=round(N/sum(N)*100, 0)][]

ever_inc_dt
```
```{r}
ever_inc_dt[, colSums(.SD), .SDcols = 4:5]
```

Mean number of incarcerations at last tick:

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_incarcerations" = mean(n_incarcerations),
           "min_n_incarcerations" = min(n_incarcerations),
           "max_n_incarcerations" = max(n_incarcerations),
           "median_n_incarcerations" = median(n_incarcerations),
           "n_ever_incarcerated_agents"=sum(n_incarcerations > 0),
         .N)] 
```   
# Incarceration Duration

Compute statistics on agents whose last release date is 
after their last incarceration date.

```{r}
agent_dt[tick == last_tick & last_release_tick >= last_incarceration_tick, 
         .(
           "mean_duration" = mean(last_release_tick - last_incarceration_tick),
           "max_duration" = max(last_release_tick - last_incarceration_tick),
           "min_duration" = min(last_release_tick - last_incarceration_tick),
           "median_duration" = median(last_release_tick - last_incarceration_tick),
         .N)] 
```



# Release Statistics

## Number

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_releases" = mean(n_releases),
           "min_n_releases" = min(n_releases),
           "max_n_releases" = max(n_releases),
           "median_n_releases" = median(n_releases),
           "n_ever_released_agents"=sum(n_releases > 0),
         .N)] 
```
## Duration
Compute statistics on agents whose last release date is 
\*before\* their last incarceration date.

```{r}
agent_dt[tick == last_tick & last_release_tick < last_incarceration_tick, 
         .(
           "mean_rel_dur" = mean(abs(last_release_tick - last_incarceration_tick)),
           "max_rel_dur" = max(abs(last_release_tick - last_incarceration_tick)),
           "min_rel_dur" = min(abs(last_release_tick - last_incarceration_tick)),
           "median_rel_dur" = median(abs(last_release_tick - last_incarceration_tick)),
         .N)] 
```


# Smoking transition behavior

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_smkg_trans" = mean(n_smkg_stat_trans),
           "max_n_smkg_trans" = max(n_smkg_stat_trans),
           "min_n_smkg_trans" = min(n_smkg_stat_trans),
           "median_n_smkg_trans" = median(n_smkg_stat_trans),
         .N)] 
```

Smoking State Distributions:

```{r}
ans <- agent_dt[tick == last_tick, .N, by=c("race", "smoking_status")][order(race)]
ans 

ans[, prop := N / sum(N), by = race][]

ans_num_smokers <- ans[smoking_status %in% c("Former", "Current", "Never"),
                       .(num_smokers = sum(N)),
                       by = race]
ans_num_smokers
sum(ans_num_smokers$num_smokers)

sim_prop_smokers_by_race <- ans_num_smokers$num_smokers/sum(ans_num_smokers$num_smokers)


```


# Alcohol transition behavior

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_alc_trans" = mean(n_alc_use_stat_trans),
           "max_n_alc_trans" = max(n_alc_use_stat_trans),
           "min_n_alc_trans" = min(n_alc_use_stat_trans),
           "median_n_alc_trans" = median(n_alc_use_stat_trans),
         .N)] 
```

Alcohol Use State Distributions:

```{r}
ans <- agent_dt[tick == last_tick, .N, by=alc_use_status]
prop <- ans$N/sum(ans$N)
cbind(ans[,1], prop)
```

# Smoking and Alcohol Use among Incarcerated Persons vs Not-

## Smoking

Proportion of current smokers among persons who have been released 0 times.

```{r}
agent_dt[tick == last_tick, .N, by = n_releases]
agent_dt[tick == last_tick, .N, by = smoking_status]

current_smoking_by_release_dcast <- dcast(agent_dt[tick == last_tick], smoking_status ~ n_releases, value.var = "id", fun.aggregate = length)

## Proportion of current smokers among persons who have been released 0 times:
current_smoking_by_release_dcast[smoking_status == "Current", sum(`0`)] / sum(current_smoking_by_release_dcast$`0`)


```


Proportion of current smokers among persons who have been released at least once:

```{r}
n_smokers_release_once <- apply(current_smoking_by_release_dcast[,c(-1, -2)], 1, sum)
n_smokers_release_once[1]/sum(n_smokers_release_once)
```

What proportion of current or former smokers are currently smoking in the
group where n_releases = 0 vs. the group n_releases > 0?

```{r}
n_releases_0 <- current_smoking_by_release_dcast[,2]
n_releases_ge1 <- current_smoking_by_release_dcast[,3:6]

n_releases_0[1]/sum(n_releases_0[1:2])
sum(n_releases_ge1[1,])/sum(n_releases_ge1[1:2,])

```

## Alcohol Use

```{r}
agent_dt[tick == last_tick, .N, by = n_releases][order(n_releases)]
agent_dt[tick == last_tick, .N, by = alc_use_status][order(alc_use_status)]

alc_use_by_release_dcast <- dcast(agent_dt[tick==last_tick,], 
                                          alc_use_status ~ n_releases, 
                                          value.var = "id", 
                                          fun.aggregate = length)
alc_use_by_release_dcast
```

What is the proportion of persons with alcohol use disorder in the
`n_releases = 0` group?

```{r}
alc_use_by_release_dcast[4,2]/sum(alc_use_by_release_dcast[,2])
```
What is the proportion of persons with alcohol use disorder in the
`n_releases > 0` group?

```{r}
alc_use_by_release_ge1 <- alc_use_by_release_dcast[,c(-1,-2)]
sum(alc_use_by_release_ge1[,4])/sum(alc_use_by_release_ge1)
```

