# Population Overview

```{r, echo=FALSE}
library(yaml)
params <- read_yaml("../../cadre/python/myparams/model_params.yaml")
#print(params)
```

A synthetic population consisting of `r params$N_AGENTS` agents between 18 and 64 years of age is simulated in an agent-based model (ABM) with social network structure. This simulated population is broadly representative of the 18-64 year old population of RI. The population size is deliberately kept small in the early stages to make the model more efficient in terms of computation time.

Each individual is assigned a set of demographic characteristics (age, race, sex), behavioral profiles (smoking, alcohol use), incarceration history. Feedback loops between the impacts of incarceration on smoking and alcohol use are simulated in the criminal-justice involved persons, and their social networks.

## Demographics

### Age

```{r, results='hide', echo=FALSE, message=FALSE, fig.show='hide'}
NAGENTS <- params$N_AGENTS
MIN_AGE <- params$MIN_AGE
MAX_AGE <- params$MAX_AGE

runif(NAGENTS, MIN_AGE, MAX_AGE)
```

Per [census data](https://censusreporter.org/profiles/04000US44-rhode-island/), 63% of the RI population is in the range of our interest (`r MIN_AGE`-`r MAX_AGE` years). We consider the population to be uniformly distributed across this range, which might not be unreasonable given the shape of the distribution above.

### Sex

```{r, results='hide', echo=FALSE, message=FALSE}
# sex (https://www.census.gov/quickfacts/RI)
FEMALE_PROP <- params$FEMALE_PROP
```

Per [census data](see%20https://www.census.gov/quickfacts/RI), the proportion of females in RI is `r FEMALE_PROP*100`%. We dichotomize `r FEMALE_PROP*100`% of our our agents as being female, and `r 100-FEMALE_PROP*100`% as being male.

### Race/Ethnicity

```{r, results='hide', echo=FALSE, message=FALSE}
WHITE <- params$RACE_DISTRIBUTION$White #not Hispanic or Latino
BLACK <- params$RACE_DISTRIBUTION$Black
HISPANIC <- params$RACE_DISTRIBUTION$Hispanic
ASIAN <- params$RACE_DISTRIBUTION$Asian
OTHER <- 1- sum(c(WHITE, BLACK, HISPANIC, ASIAN))
```

Per [census data](see%20https://www.census.gov/quickfacts/RI), we parameterize the population as belonging to 5 race/ethnic categories: White, Black, Hispanic, Asian, and Other.

## Substance Use Behaviors

### Smoking

#### Behavioral States

```{r, results='show', echo=FALSE, message=FALSE}
SMOKING_PREV <-params$SMOKING_PREV
```

Per the [CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm), the national prevalence of current smokers in the US is `r SMOKING_PREV*100`%. ( The CDC defines "a current smoker" as someone who "who reported smoking at least 100 cigarettes during their lifetime and who, at the time they participated in a survey about this topic, reported smoking every day or some days.")

The [CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm) also stratifies the national smoking rates by age, sex, and race/ethnicity. There is also a geographic prevalence of smoking (about 10% in the northeastern US), that is not stratified by demographic variables.

#### Transitions between Behavioral States

### Alcohol Use

#### Behavioral States

```{r, results='hide', echo=FALSE, message=FALSE}
ABSTAINERS <- params$ALC_USE_PROPS$A
OCCASIONAL <- params$ALC_USE_PROPS$O
REGULAR <- params$ALC_USE_PROPS$R
AUD <- params$ALC_USE_PROPS$D
```

We classify alcohol use behaviors in one of four categories, defined by frequency of use, as in a previous pair of modeling studies @Apostolopoulos2018I; @Apostolopoulos2018II. The four categories are: "abstainers", "occassional users", "regular users", "persons with high-risk use or Alcohol Use Disorder (AUD)".

To parameterize the distribution of persons in each of these categories, data from the National Comorbidity Survey Replication, as published in @Kalaydjian2009, where:

    * (A) 91.7% (SE=0.9) reported >1 sip of alcohol at some time in their life, 
    * (B) 72.9% (1.3) reported using alcohol regularly at some time in their life
    * (C) 13.2% (0.6) met criteria for alcohol abuse at some time in their life, and 
    * (D) 5.4% (0.3) met criteria for alcohol dependence at some time in their life.

Thus, the proportion of abstainers is $1-A$ = `r ABSTAINERS`%. Category B above is taken to be persons who use alcohol occasionally (`r OCCASIONAL`%), Category C as "regular users" (`r REGULAR`). Category D as High-Risk/AUD = 100-(A+B+C) = `r AUD`%, which is close to the value D reported by @Kalaydjian2009.

#### Transitions Between Behavioral States

## Incarceration

```{r, results='hide', echo=FALSE, message=FALSE}
# 2-week risk #macmadu 2021 (table 1)
PROB_DAILY_INCARCERATION <- params$PROBABILITY_DAILY_INCARCERATION
PROB_DAILY_RECIDIVISM <- params$PROBABILITY_DAILY_RECIDIVISM
  # we'll break down the above by age and race
```

Incarceration functions by defining two locations: 
a "correctional setting" (CS), and a "home" setting. Per @Macmadu2021, the 
2-week incarceration probability for RI persons is 
`r params$PROBABILITY_DAILY_INCARCER`%. 

