---
title: "Technical Appendix to the CADRE Model"
author: "CADRE Project 8 Team"
date: "`r Sys.Date()`"
output: pdf_document
bibliography:
- book.bib
- packages.bib
- "cadre-appendix-library.bib"
link-citations: yes
documentclass: book
---
  
# Overview

This is an early draft of what will become the companion Appendix to the modeling papers that we publish. Narrative text and the R code are interwoven. The computer code is also available in a public [GitHub repository](https://github.com/khanna7/cadre).

This document is written as a **bookdown** document in **Markdown** to facilitate [reproducibility](https://bookdown.org/m_p_sperrin/intro_to_r_2020/projects-and-reproducibility.html). One advantage of this format is that the data sources, decisions about model parameterization, and the computer code implementing these decisions all exist in one place. This makes it easier to understand the reasoning behind the decisions we make, and appropriately source, cite, or derive our data as we go.

The document can also be automatically updated whenever the underlying code or data change.

## Usage Notes

Each **bookdown** chapter is an .Rmd file, and each .Rmd file can contain one (and only one) chapter. A chapter *must* start with a first-level heading: `# A good chapter`, and can contain one (and only one) first-level heading.

Use second-level and higher headings within chapters like: `## A short section` or `### An even shorter section`.

The `index.Rmd` file is required, and is also our first book chapter. This creates the homepage when the book is "render"ed.

## Render book

You can render the HTML version of this example book without changing anything:

1.  Find the **Build** pane in the RStudio IDE, and

2.  Click on **Build Book**, then select your output format, or select "All formats" if you'd like to use multiple formats from the same book source files.

Or build the book from the R console:

```{r, eval=FALSE}
bookdown::render_book()
```

To render this example to PDF as a `bookdown::pdf_book`, you'll need to install XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

## Viewing Instructions

As you work, you may start a local server to live preview this HTML book. This preview will update as you edit the book when you save individual .Rmd files. You can start the server in a work session by using the RStudio add-in "Preview book", or from the R console:

```{r eval=FALSE}
bookdown::serve_book()
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

# Population Overview

```{r, echo=FALSE}
library(yaml)
filename <- "/Volumes/GoogleDrive/My Drive/code/cadre/python/myparams/model_params.yaml"
params <- read_yaml(filename)

```

A synthetic population consisting of `r params$N_AGENTS` agents between 18 and 64 years of age is simulated in an agent-based model (ABM) with social network structure. This simulated population is broadly representative of the 18-64 year old population of RI. The population size is deliberately kept relatively small in the early stages to make the model more efficient in terms of computation time.

Each individual is assigned a set of demographic characteristics (age, race, sex), behavioral profiles (smoking, alcohol use), incarceration history. Feedback loops between the impacts of incarceration on smoking and alcohol use are simulated in the criminal-justice involved persons, and their social networks.

## Demographics


```{r, results='hide', echo=FALSE, message=FALSE, fig.show='hide'}
NAGENTS <- params$N_AGENTS
MIN_AGE <- params$MIN_AGE
MAX_AGE <- params$MAX_AGE

#runif(NAGENTS, MIN_AGE, MAX_AGE)
```

Per [census data](https://censusreporter.org/profiles/04000US44-rhode-island/), 63% of the RI population is in the range of our interest (`r MIN_AGE`-`r MAX_AGE` years). We consider the population to be uniformly distributed across this range, which might not be unreasonable given the shape of the distribution above.



```{r, results='hide', echo=FALSE, message=FALSE}
# sex (https://www.census.gov/quickfacts/RI)
FEMALE_PROP <- params$FEMALE_PROP
```

Per [census data](see%20https://www.census.gov/quickfacts/RI), the proportion of females in RI is `r FEMALE_PROP*100`%. We dichotomize `r FEMALE_PROP*100`% of our our agents as being female, and `r 100-FEMALE_PROP*100`% as being male.



```{r, results='hide', echo=FALSE, message=FALSE}
WHITE <- params$RACE_DISTRIBUTION$White #not Hispanic or Latino
BLACK <- params$RACE_DISTRIBUTION$Black
HISPANIC <- params$RACE_DISTRIBUTION$Hispanic
ASIAN <- params$RACE_DISTRIBUTION$Asian
OTHER <- 1- sum(c(WHITE, BLACK, HISPANIC, ASIAN))
```

Per [census data](see%20https://www.census.gov/quickfacts/RI), we parameterize the population as belonging to 5 race/ethnic categories: White, Black, Hispanic, Asian, and Other.

# Substance Use Behaviors 

## Alcohol Use: Behavioral States

```{r, results='hide', echo=FALSE, message=FALSE}
ABSTAINERS <- params$ALC_USE_PROPS$A
OCCASIONAL <- params$ALC_USE_PROPS$O
REGULAR <- params$ALC_USE_PROPS$R
AUD <- params$ALC_USE_PROPS$D
```

We classify alcohol use behaviors in one of four categories, defined by frequency 
of use, as in a previous pair of 
modeling studies @Apostolopoulos2018I; @Apostolopoulos2018II. 
The four categories are: "abstainers", "occassional users", "regular users", 
"persons with high-risk use or Alcohol Use Disorder (AUD)".

To parameterize the distribution of persons in each of these categories, 
data from the National Comorbidity Survey Replication, 
as published in @Kalaydjian2009, where:

    * (A) 91.7% (SE=0.9) reported >1 sip of alcohol at some time in their life, 
    * (B) 72.9% (1.3) reported using alcohol regularly at some time in their life
    * (C) 13.2% (0.6) met criteria for alcohol abuse at some time in their life, and 
    * (D) 5.4% (0.3) met criteria for alcohol dependence at some time in their life.

Thus, the proportion of abstainers is $1-A$ = `r ABSTAINERS*100`%. 
Category B above is taken to be persons who use alcohol occasionally 
(`r OCCASIONAL*100`%), Category C as "regular users" (`r REGULAR*100`), and 
Category D denotes persons in the High-Risk/AUD category = 100-(A+B+C) = 
`r AUD*100`%, which is very close to the value D reported by @Kalaydjian2009.

## Alcohol Use: State Transitions

We simulate transitions between states of alcohol use, thus incorporating changes
in people's drinking behavior over the course of their lives. Because the rates
of transitions between the four states of alcohol were not 
available, we used the Metropolis Hastings algorithm to compute these transition 
rates given the distribution of persons in each of the states as defined above.  
The algorithm allows us to derive the transition rates so that the
distribution of persons in each state of alcohol use remain consistent with the
data presented above (i.e., the distribution remains ``stationary''). 

Using the commonly used notation $\pi$ to denote the stationary distribution 
as a row vector, we have

\begin{equation} 
  \pi = (`r ABSTAINERS`,  
  `r OCCASIONAL`, 
  `r REGULAR`,  
  `r AUD`)
\end{equation} 

We additionally set up rules for this system that permit 
transitions between the four states of alcohol use, 
represented as a transition matrix $\mathbf{P_{ALCOHOL}}$ below:

\begin{equation} 
  \mathbf{P_{ALCOHOL}} = 
\begin{pmatrix}
  p_{11} & p_{12} & p_{13} & p_{14} \\
  p_{21} & p_{22} & p_{23} & p_{24} \\
  p_{31} & p_{32} & p_{33} & p_{34} \\
  p_{41} & p_{42} & p_{43} & p_{44} \\
 \end{pmatrix}.
\end{equation} 

The numbers 1, 2, 3, and 4 respectively denote states of 
abstention, occasional use, regular use, and heavy drinking/AUD respectively.
The probabilities denoted by $p_{xy}$ denote the probability of tranisitioning
between states $x$ and $y$ in a given day. Our system prohibits transitions by 
that would involve changing two or more levels in a day. Thus,  

\begin{equation} 
  \mathbf{P_{ALCOHOL}} = 
\begin{pmatrix}
  p_{11} & p_{12} & 0 & 0 \\
  p_{21} & p_{22} & p_{23} & 0 \\
  0 & p_{32} & p_{33} & p_{34} \\
  0 & 0 & p_{43} & p_{44} \\
 \end{pmatrix}.
\end{equation} 

The non-zero entries in a given row in $\mathbf{P_{ALCOHOL}}$
describe all the possible transitions that any agent can make in a given day, 
and therefore sum to 1.

Denoting the unknown transition rates as a row vector 
$ \mathbf{\lambda} = (\lambda_1, \lambda_2, \lambda_3, \lambda_4), $
we use the Metropoplis Hastings algorithm to solve the following system of 
equations:

\begin{equation} 
\mathbf{\lambda} \mathbf{P_{ALCOHOL}} = \mathbf{\pi}.
\end{equation} 

Solving the above [system of equations](https://drive.google.com/file/d/1bnFgEiRK4vFikyscSvq8-W7vRoMZecoc/view?usp=sharing) 
yields:

\begin{matrix}
  83 p_{12} & = & 729 p_{21} \\
  729 p_{23} &=& 132 p_{32} \\
  132 p_{34} &=& 56 p_{43} \\
  \end{matrix}.

Any set of probability values that match the above criteria would provide a 
solution to the above set of equations. We choose a set that gives considers
high probabilities for people to remain in their alcohol use state, and relatively
low probabilities to transition out of it:

```{r, results='hide', echo=FALSE, message=FALSE, fig.show='hide'}

TRANS_PROB_0_1 <- params$ALC_USE_STATES$TRANS_PROB_0_1
TRANS_PROB_1_2 <- params$ALC_USE_STATES$TRANS_PROB_1_2
TRANS_PROB_2_3 <- params$ALC_USE_STATES$TRANS_PROB_2_3

TRANS_PROB_1_0 <- params$ALC_USE_STATES$TRANS_PROB_1_0
TRANS_PROB_2_1 <- params$ALC_USE_STATES$TRANS_PROB_2_1
TRANS_PROB_3_2 <- params$ALC_USE_STATES$TRANS_PROB_3_2

```

\begin{equation} 
  \mathbf{P_{ALCOHOL}} = 
\begin{pmatrix}
  `r 1-TRANS_PROB_0_1` & `r TRANS_PROB_0_1` & 0 & 0 \\
  `r TRANS_PROB_1_0`  & `r 1-(TRANS_PROB_1_0 + TRANS_PROB_1_2)` & `r TRANS_PROB_1_2` & 0 \\
  0 & `r TRANS_PROB_2_1` & `r  1-(TRANS_PROB_2_1 + TRANS_PROB_2_3)` & `r TRANS_PROB_2_3` \\
  0 & 0 & `r TRANS_PROB_3_2`  & `r 1-TRANS_PROB_3_2` \\
 \end{pmatrix}.
\end{equation}

## Smoking: Behavioral States

Per the 
[CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm),
and as published in previous studies (@Levy2005), 
smoking behavior is often classified in one of three states:
"Never Smokers" (i.e., persons who have smoked less than 100 cigarettes in their lifetime),
"Current Smokers" (i.e., persons who have smoked at least 100 cigarettes 
in their lifetime and at least once recently),
and, 
"Former Smokers" (i.e., persons who have smoked at least 100 cigarettes 
in their lifetime but none recently). 
For current and former smokers, we define recency as 30 days, as has been 
done previously (@Ahmad2005), to formulate a state chart as shown below: 

```{r, echo=FALSE}
library(DiagrammeR)
mermaid("
graph LR
  A(Never Smoker) --> B(Current Smoker)
  B --> C(Former Smoker)
  C --> B
")
```

```{r, results='hide', echo=FALSE, message=FALSE}
SMOKING_PREV <-params$SMOKING_PREV

SMOKING_PREV_TAB <- as.data.frame(matrix(c(
  "White Men", SMOKING_PREV$WHITE_MALE_NEVER, SMOKING_PREV$WHITE_MALE_CURRENT, SMOKING_PREV$WHITE_MALE_FORMER,
  "White Women", SMOKING_PREV$WHITE_FEMALE_NEVER, SMOKING_PREV$WHITE_FEMALE_CURRENT, SMOKING_PREV$WHITE_FEMALE_FORMER, 
    "Black Men", SMOKING_PREV$BLACK_MALE_NEVER, SMOKING_PREV$BLACK_MALE_CURRENT, SMOKING_PREV$BLACK_MALE_FORMER,
    "Black Women", SMOKING_PREV$BLACK_FEMALE_NEVER, SMOKING_PREV$BLACK_FEMALE_CURRENT, SMOKING_PREV$BLACK_FEMALE_FORMER,
    "Hispanic Men", SMOKING_PREV$HISPANIC_MALE_NEVER, SMOKING_PREV$HISPANIC_MALE_CURRENT, SMOKING_PREV$HISPANIC_MALE_FORMER,
  "Hispanic Women", SMOKING_PREV$HISPANIC_FEMALE_NEVER, SMOKING_PREV$HISPANIC_FEMALE_CURRENT, SMOKING_PREV$HISPANIC_FEMALE_FORMER,
  "Asian Men", SMOKING_PREV$ASIAN_MALE_NEVER, SMOKING_PREV$ASIAN_MALE_CURRENT, SMOKING_PREV$ASIAN_MALE_FORMER,
  "Asian Women", SMOKING_PREV$ASIAN_FEMALE_NEVER, SMOKING_PREV$ASIAN_FEMALE_CURRENT, SMOKING_PREV$ASIAN_FEMALE_FORMER,
    "Other Men", SMOKING_PREV$OTHER_MALE_NEVER, SMOKING_PREV$OTHER_MALE_CURRENT, SMOKING_PREV$OTHER_MALE_FORMER,
  "Other Women", SMOKING_PREV$OTHER_FEMALE_NEVER, SMOKING_PREV$OTHER_FEMALE_CURRENT, SMOKING_PREV$OTHER_FEMALE_FORMER
),
  byrow = TRUE, ncol = 4)
)
colnames(SMOKING_PREV_TAB) <- c("Population", "Never", "Current", "Former")
SMOKING_PREV_TAB
```

```{r, echo=FALSE}
knitr::kable(SMOKING_PREV_TAB, "pipe", caption = "Smoking States by Race/Ethnicity")
```
 
Per the [CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm), the national prevalence of current smokers in the US is 
( The CDC defines "a current smoker" as someone who "who reported smoking at least 100 cigarettes during their lifetime and who, at the time they participated in a survey about this topic, reported smoking every day or some days.")

The [CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm) also stratifies the national smoking rates by age, sex, and race/ethnicity. There is also a geographic prevalence of smoking (about 10% in the northeastern US), that is not stratified by demographic variables.



```{r, results='hide', echo=FALSE, message=FALSE}
params["SMOKING_PREV"]
```


## Smoking: State Transitions


# Incarceration and Recidivism

```{r, results='hide', echo=FALSE, message=FALSE}
# 2-week risk #macmadu 2021 (table 1)
PROB_DAILY_INCARCERATION <- params$PROBABILITY_DAILY_INCARCERATION
PROB_DAILY_RECIDIVISM <- params$PROBABILITY_DAILY_RECIDIVISM
  # we'll break down the above by age and race
```

Incarceration functions by defining two locations: a "correctional setting" (CS), and a "home" setting. Per @Macmadu2021, the 2-week incarceration probability for RI persons is `r 1.7`%.

Considering that the risk of incarceration is distributed identically and independently across the biweekly period (i.e., 14 days), we obtain the daily incarceration rate using the binomial formula:

$$1 - (1-p)^{14} = 0.017$$

where p is the daily incarceration probability; (1-p) is the probability that a person does not get incarcerated in one day; $(1-p)^{14}$ is the probability that the person does \*not\* get incarcerated across the 14 days. Subtracting this probability from 1 gives the probability that the person will be incarcerated across the 14 days.

Therefore,

$$(1-p)^{14} = 1-0.017 = 0.983.$$ Therefore, $$(1-p) = 0.9831^{1/14} = 0.99877$$ and we conclude that the daily incarceration probability $p$ is `r 1-((1-0.017)^(1/14))`.

Similarly per @Macmadu2021 et al, the 3-year recidivism probability for men and women with prior incarceration history is 51% and 36% respectively. For men, the daily recidivism probability ($p_{RM}$) can be obtained by solving:

$$1 - (1-p_{RM})^{365 \times 3} = 0.51.$$ Thus, $$(1-p_{RM})^{365 \times 3} = 0.49.$$ 
Therefore, 
$p_{RM}$ = `r 1 - ((0.49)^(1/(365*3)))`.

Similarly, for women,
$$1 - (1-p_{RF})^{365\times3} = 0.36$$
where $p_{RF}$ is the daily probability of recidivism for women. 
We solve this equation to get 
$ p_{RF} =$ 
`r 1 - ((0.64)^(1/1095))`. 

(There is a bit of an anomaly here since the daily probability of recividism 
for men and women is lower than the daily probability of first-time incarceration.
I emailed Dr. Macmadu to talk about this but I didn't hear back. 
We'll probably want to talk about alternate parameterizations.)

# Post-Release Smoking and Alcohol Use

# Social Network Structure 

[text here]


# Bibliography



