---
title: "Agent Outome Analysis"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


We confirm below various attributes of the agents that are logged.

We will do this in `data.table` for speed. 

```{r setup, include=FALSE}
rm(list=ls())
renv::load()
library(ggplot2)
library(data.table)
library(yaml)
library(dplyr)
```

```{r}
agent_dt <- 
  fread("/oscar/data/akhann16/akhann16/cadre_simulated_data/emews_experiments/test-alt-loc/instance_1/output/agent_log.csv")
str(agent_dt)
glimpse(agent_dt)
last_tick <- max(agent_dt$tick)
print(last_tick)
```

## Outcome Variables

### Smoking States as a Function of Time Since Release

We want to compute the distribution of smoking states among the set of
agents who have been released for 1 day, and at most 1 week, 
2 weeks, 1 month, 6 months, 
and 1 year. 


- Distribution of persons who have been released 1 day:  

```{r}
# Create a new column indicating if an agent was released the previous day
agent_dt[, released_1_day_ago := (tick - last_release_tick == 1)]

# For each tick, summarize the distribution of smoking states for agents released the previous day
distribution_1_day <- agent_dt[released_1_day_ago == TRUE, 
                               .(count = .N), 
                               by = .(tick, smoking_status)]


# Calculate the total count for each smoking status
total_counts <- distribution_1_day[, .(total_count = sum(count)), by = smoking_status]

# Calculate the overall total
overall_total <- sum(total_counts$total_count)

# Calculate the percentage for each smoking status
total_counts[, percentage := (total_count / overall_total) * 100]

# View the percentages
total_counts[, .(smoking_status, percentage)]

```

- Distribution of persons who have been released at most 6 months: 

```{r}
# Create a new column indicating if an agent was released 6 months or less ago
agent_dt[, released_6_months_or_less := (tick - last_release_tick <= 180) & (last_release_tick != -1)]

# Sum the count of agents who have been released 6 months or less ago
count_6_months_or_less <- agent_dt[released_6_months_or_less == TRUE, .N]

# Filter the data for agents who were released 6 months or less ago and group by their smoking status
smoking_distribution_6_months_or_less <- 
  agent_dt[released_6_months_or_less == TRUE, .N, by = smoking_status]

# Rename the count column
setnames(smoking_distribution_6_months_or_less, "N", "count")

# Calculate the percentage for each smoking status within this subset
smoking_distribution_6_months_or_less[, percentage := (count / sum(count)) * 100]

# View the result
smoking_distribution_6_months_or_less
```

- Distribution of persons who have been released at most 1 year:

```{r}
# Create a new column indicating if an agent was released 6 months or less ago
agent_dt[, released_1_year_or_less := (tick - last_release_tick <= 365) & 
           (last_release_tick != -1)]

# Sum the count of agents who have been released 6 months or less ago
count_1_year_or_less <- agent_dt[released_1_year_or_less == TRUE, .N]

# Filter the data for agents who were released 6 months or less ago and group by their smoking status
smoking_distribution_1_year_or_less <- 
  agent_dt[released_1_year_or_less == TRUE, .N, by = smoking_status]

# Rename the count column
setnames(smoking_distribution_1_year_or_less, "N", "count")

# Calculate the percentage for each smoking status within this subset
smoking_distribution_1_year_or_less[, percentage := (count / sum(count)) * 100]

# View the result
smoking_distribution_1_year_or_less[]
```

