---
title: "R Notebook"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


We confirm below various attributes of the agents that are logged.

We will do this in `data.table` for speed. 

```{r setup, include=FALSE}
rm(list=ls())
renv::load()
library(ggplot2)
library(data.table)
library(yaml)
```

Read input parameters:¬

```{r}
input_params <- read_yaml("~/code/cadre/python/myparams/model_params.yaml")
```


Read the agent data into data-table:

```{r}
agent_dt <- fread("~/code/cadre/python/output/agent_log_9.csv")
str(agent_dt)
last_tick <- max(agent_dt$tick)
print(last_tick)
```

How many unique agents are in the dataset?

```{r}
length(unique(agent_dt$id))
range(unique(agent_dt$id))
```

There are `r length(unique(agent_dt$id))` unique agents in the dataset. Their IDs 
range from `r min((unique(agent_dt$id)))` to `r max((unique(agent_dt$id)))`.

# Demographics

```{r}
agent_dt_by_race <- agent_dt[tick == last_tick, .N, by=race][,"per_cent":=round(N/sum(N)*100)][order(race)]
target_values <- data.frame(race = names(input_params$RACE_DISTRIBUTION), 
                            target_pct = unlist(input_params$RACE_DISTRIBUTION))
```


Plot race distribution over time:

```{r}
# Define the time ticks you want to analyze
time_ticks <- c(seq(1, last_tick, by = 10), last_tick)

# Calculate the proportion of each race at the specified time ticks
race_proportions_by_tick <- agent_dt[tick %in% time_ticks, .(
  White = sum(race == "White") / .N,
  Black = sum(race == "Black") / .N,
  Hispanic = sum(race == "Hispanic") / .N,
  Asian = sum(race == "Asian") / .N
), by = tick]

# Convert the data from wide to long format
race_proportions_long <- melt(race_proportions_by_tick, id.vars = "tick", variable.name = "race", value.name = "proportion")

# Create a time series plot of race proportions
# Create a time series plot of race proportions
ggplot(race_proportions_long, aes(x = tick, y = proportion, color = race, group = race)) +
  geom_line() +
  labs(title = "Agent Race Proportions Over Time",
       x = "Time (Days)",
       y = "Race Distributions") +
  scale_color_manual(values = c("#377eb8", "#ff7f00", "#4daf4a", "#e41a1c")) +
  theme_minimal() +
  theme(legend.title = element_blank())+
  scale_y_continuous(limits = c(0, 0.8), breaks = seq(0, 0.8, by = 0.1))
```

```{r echo=TRUE}
agent_dt[tick==last_tick, .N, by=female][,"%":=N/sum(N)][order(female)]
female_target <- input_params$FEMALE_PROP

gender_pct <- agent_dt[tick == last_tick, .N, by = female][, "%":= N/sum(N)][order(female)]
```
```{r}
# calculate percentages
gender_pct <- agent_dt[tick == last_tick, .N, by = female][, "%":= N/sum(N)][order(female)]

# read in YAML file with target value
female_target <- input_params$FEMALE_PROP

# calculate target percentages for each gender
female_target_pct <- female_target
male_target_pct <- 1 - female_target

# calculate actual percentages for each gender
female_actual_pct <- gender_pct[female == 1, `%`]
male_actual_pct <- gender_pct[female == 0, `%`]


# create plot
ggplot(gender_pct, aes(x = ifelse(female == 0, "Male", "Female"), y = `%`, fill = ifelse(female == 0, "Male", "Female"))) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02")) +
  annotate("text", x = 1, y = female_actual_pct, 
           label = paste0("Female Target = ", round(female_actual_pct*100, 1), "%"), 
           color = "#1b9e77", size = 4, vjust = 0) +
  annotate("text", x = 2, y = male_actual_pct, 
           label = paste0("Male Target = ", round(male_actual_pct*100, 1), "%"), 
           color = "#d95f02", size = 4, vjust = 0) +
  labs(title = "",
       x = "Time (Days)",
       y = "Sex Distributions",
       fill = "") +
  theme_minimal()




```

Sex distributions over time:

```{r}
# Define the time ticks you want to analyze
time_ticks <- c(seq(1, last_tick, by = 10), last_tick)

# Calculate the proportion of males and females at the specified time ticks
gender_proportions_by_tick <- agent_dt[tick %in% time_ticks, .(
  male = sum(female == 0) / .N,
  female = sum(female == 1) / .N
), by = tick]

# Convert the data from wide to long format
gender_proportions_long <- melt(gender_proportions_by_tick, id.vars = "tick", variable.name = "gender", value.name = "proportion")

# Create a time series plot of male and female proportions
ggplot(gender_proportions_long, aes(x = tick, y = proportion, color = gender, group = gender)) +
  geom_line() +
  labs(title = "",
       x = "Time (Days)",
       y = "Sex Distributions") +
  scale_color_manual(values = c("#1b9e77", "#d95f02"), labels = c("Female", "Male")) +
  theme_minimal() +
  theme(legend.title = element_blank())+
  scale_y_continuous(limits = c(0.4, 0.6), breaks = seq(0.4, 0.6, by = 0.1))

```


```{r}
agebreaks <- c(18, 25, 35, 45, 55, 65)
agelabels <- c("18-24", "25-34", "35-44", "45-54", "55-64")
```


```{r}
agent_dt[tick == last_tick, .N, ]
setDT(agent_dt)[ , age_groups := cut(age, 
                                breaks = agebreaks, 
                                include.lowest = TRUE,
                                right = FALSE, 
                                labels = agelabels)]
nrow(agent_dt[tick == last_tick])
```

```{r}
agent_dt[tick == last_tick, .N, by=c("age_groups", "race", "female")][order(age_groups, race, female)]
agent_dt[tick == last_tick, .N, by=c("race", "age_groups", "female")][order(race, age_groups)]
```
```{r}
agent_dt[tick==last_tick, .N, by=c("race", "female")][,"%":=N/sum(N)*100][order(race)]
```

```{r}
agent_dt[tick==last_tick, .N, by=c("age_groups")][,"%":=round(N/sum(N)*100)][order(age_groups)]
```
Median age at the start: 

```{r}
agent_dt[tick==1, median(age)]
```

Median age at the end:

```{r}
nrow(agent_dt[tick==last_tick])
agent_dt[tick==last_tick, median(age)]
```

```{r}

# Create age groups
agent_dt_by_age_group <- agent_dt[tick==last_tick, .N, by = .(age_group = cut(age, breaks = c(18, 25, 35, 45, 55, 65, 75, 80, 84)))]

# Create bar chart
ggplot(agent_dt_by_age_group, aes(x = age_group, y = N, fill = age_group)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00", "#ffff33", "#a65628", "#f781bf")) +
  labs(title = "Agent Age Distribution",
       x = "Age Group",
       y = "Count",
       fill = "Age Group") +
  theme_minimal()

```

# Baseline Smoking Behaviors


At the first tick:

```{r}
agent_dt[tick==1 & race == "Black" & female == 0, .N, by=c("smoking_status")][]

agent_dt[tick==1 & race == "Black" & female == 0, .N, by=c("smoking_status")][,"sim_prop":= round(N/sum(N), 3)][]


# targets
BLACK_MALE_CURRENT = input_params$SMOKING_PREV$BLACK_MALE_CURRENT
BLACK_MALE_FORMER = input_params$SMOKING_PREV$BLACK_MALE_FORMER
BLACK_MALE_NEVER = input_params$SMOKING_PREV$BLACK_MALE_NEVER

smkg_prev_black_men_target <- c(BLACK_MALE_CURRENT, BLACK_MALE_FORMER, BLACK_MALE_NEVER)

smkg_prev_black_men_target

## Add difference between the two target and the sim

```

At the last tick:
```{r}
smkg_blackmen_last_tick <- agent_dt[tick==last_tick & race == "Black" & female == 0, .N, by=c("smoking_status")][,"prop":=round(N/sum(N), 3)][]
smkg_blackmen_last_tick

smkg_prev_black_men_target

```

```{r}
# Count the number and percentage of each smoking status
smoking_dt <- agent_dt[tick == last_tick, .N, by = smoking_status][, `:=` (percentage = round(N/sum(N)*100, 2))]
# Create the bar chart
ggplot(smoking_dt, aes(x = smoking_status, y = percentage, fill = smoking_status)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("#fec44f", "#43a2ca", "#7bccc4")) +
  labs(title = "Agent Smoking Status Distribution",
       x = "Smoking Status",
       y = "Percentage",
       fill = "Smoking Status") +
  theme_minimal()

```

# Baseline Alcohol Use

```{r}
alc_use_last_tick <- 
  agent_dt[tick == last_tick, .N, by = c("alc_use_status")][, 
  "%" := round(N /sum(N), 3)][
  order(alc_use_status)][
  ]

# targets
alc_use_props <- input_params$ALC_USE_PROPS
alc_use_targets <- c(alc_use_props$A, alc_use_props$O, alc_use_props$R, alc_use_props$D)

cbind(alc_use_last_tick[["%"]], alc_use_targets)
```

```{r}
# Calculate the percentages for alcohol use status
alc_use_dt <- agent_dt[tick == last_tick, .N, by = alc_use_status][, `:=` (percentage = round(N/sum(N)*100, 2))]
# Order the data table by alcohol use status
setorder(alc_use_dt, alc_use_status)
# Rename the alcohol use status codes to descriptive labels
alc_use_dt[, alc_use_status := factor(alc_use_status, levels = c(0, 1, 2, 3), 
                                      labels = c("Lifetime Abstainer", 
                                                 "Ocassional", 
                                                 "Regular", 
                                                 "Heavy"))]

ggplot(alc_use_dt, aes(x = alc_use_status, y = percentage, fill = alc_use_status)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Agent Alcohol Use Status Distribution",
       x = "Use Frequency",
       y = "Percentage",
       fill = "Alcohol Use Status") +
  theme_minimal() +
  guides(fill = "none") +
  geom_text(aes(label = paste0(as.numeric(input_params$ALC_USE_PROPS)*100, "%")), 
            position = position_stack(vjust = 1.02))+
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 10))


```

# Smoking and Alcohol Co-Use


 
## Conditional Probabilities on Co-Use

### Approach 1

Write functions to compute co-use probabilities at any  time step:

```{r}


# Functions for computing probabilities at a specific tick
compute_p_smoker_given_alc_user <- function(tick_data) {
  alc_users <- tick_data[tick_data$alc_use_status != 0,]
  n_smokers_and_alc_users <-
    sum(alc_users$smoking_status == "Current")
  p_smoker_given_alc_user <-
    n_smokers_and_alc_users / nrow(alc_users)
  return(p_smoker_given_alc_user)
}

compute_p_alc_user_given_smoker <- function(tick_data) {
  smokers <- tick_data[tick_data$smoking_status == "Current",]
  n_smokers_and_alc_users <- sum(smokers$alc_use_status != 0)
  p_alc_user_given_smoker <-
    n_smokers_and_alc_users / nrow(smokers)
  return(p_alc_user_given_smoker)
}

compute_p_abstainer_given_never_smoker <- function(tick_data) {
  never_smokers <- tick_data[tick_data$smoking_status == "Never",]
  n_abstainers_and_never_smokers <-
    sum(never_smokers$alc_use_status == 0)
  p_abstainer_given_never_smoker <-
    n_abstainers_and_never_smokers / nrow(never_smokers)
  return(p_abstainer_given_never_smoker)
}

compute_p_never_smoker_given_abstainer <- function(tick_data) {
  abstainers <- tick_data[tick_data$alc_use_status == 0,]
  n_abstainers_and_never_smokers <-
    sum(abstainers$smoking_status == "Never")
  p_never_smoker_given_abstainer <-
    n_abstainers_and_never_smokers / nrow(abstainers)
  return(p_never_smoker_given_abstainer)
}

compute_p_current_or_former_smoker_given_alc_user <-
  function(tick_data) {
    alc_users <- tick_data[tick_data$alc_use_status != 0,]
    n_current_or_former_smokers_and_alc_users <-
      sum(alc_users$smoking_status %in% c("Current", "Former"))
    p_current_or_former_smoker_given_alc_user <-
      n_current_or_former_smokers_and_alc_users / nrow(alc_users)
    return(p_current_or_former_smoker_given_alc_user)
  }

compute_p_alc_user_given_current_or_former_smoker <-
  function(tick_data) {
    current_or_former_smokers <-
      tick_data[tick_data$smoking_status %in% c("Current", "Former"),]
    n_current_or_former_smokers_and_alc_users <-
      sum(current_or_former_smokers$alc_use_status != 0)
    p_alc_user_given_current_or_former_smoker <-
      n_current_or_former_smokers_and_alc_users / nrow(current_or_former_smokers)
    return(p_alc_user_given_current_or_former_smoker)
  }

```

Apply these functions to the first and last tick data:

```{r}
first_tick_data <- agent_dt[agent_dt$tick == 1,]
last_tick_data <- agent_dt[agent_dt$tick == max(tick)]
```

```{r}
compute_p_smoker_given_alc_user(first_tick_data)
compute_p_alc_user_given_smoker(first_tick_data)
compute_p_abstainer_given_never_smoker(first_tick_data)
compute_p_never_smoker_given_abstainer(first_tick_data)
compute_p_current_or_former_smoker_given_alc_user(first_tick_data)
compute_p_alc_user_given_current_or_former_smoker(first_tick_data)
```



```{r}
compute_p_smoker_given_alc_user(last_tick_data)
compute_p_alc_user_given_smoker(last_tick_data)
compute_p_abstainer_given_never_smoker(last_tick_data)
compute_p_never_smoker_given_abstainer(last_tick_data)
compute_p_current_or_former_smoker_given_alc_user(last_tick_data)
compute_p_alc_user_given_current_or_former_smoker(last_tick_data)
```


```{r}
# Create a contingency table of smoking and alcohol use status
smoking_alc_use_table <- dcast(agent_dt[tick == last_tick], smoking_status ~ alc_use_status, value.var = "id", fun.aggregate = length)
smoking_alc_use_table <- smoking_alc_use_table[, .(smoking_status, `Lifetime Abstainer` = `0`, Occasional = `1`, Regular = `2`, Heavy = `3`)]
smoking_alc_use_melt <- melt(smoking_alc_use_table, id.vars = "smoking_status",
                             variable.name = "alcohol_use")
```

### Approach II

Create a contingency table at the last tick:

```{r}
smkg_alc_use_xtab <- dcast(smoking_alc_use_melt, smoking_status ~ alcohol_use, value.var = "value")
smkg_alc_use_xtab

smoking_alc_col_props <- prop.table(as.matrix(smkg_alc_use_xtab[,-1]), margin = 1)
smoking_alc_row_props <- prop.table(as.matrix(smkg_alc_use_xtab[,-1]), margin = 2)

smoking_alc_col_props <- cbind(smkg_alc_use_xtab[,1], round(smoking_alc_col_props, 3))
smoking_alc_row_props <- cbind(smkg_alc_use_xtab[,1], round(smoking_alc_row_props, 3))

print(smoking_alc_col_props)
print(smoking_alc_row_props)

rowSums(smoking_alc_col_props[,-1])
colSums(smoking_alc_row_props[,-1])
```

P(Current Smoker |Heavy Drinker) = `r (as.numeric(smoking_alc_row_props["Current","Heavy"]))`  
P(Current Smoker | Regular Drinker) = `r (as.numeric(smoking_alc_row_props["Current","Regular"]))`
P(Current Smoker | Occasional Drinker)  = `r (as.numeric(smoking_alc_row_props["Current","Occasional"]))`

Alcohol Use conditioned on Current Smoking:
P(Heavy Drinker | Current Smoker) = `r (as.numeric(smoking_alc_col_props["Current","Heavy"]))` 
P(Regular Drinker| Current Smoker) = `r (as.numeric(smoking_alc_col_props["Current","Regular"]))`  
P(Occasional Drinker | Current Smoker) = `r (as.numeric(smoking_alc_col_props["Current","Occasional"]))`  


## Marginal Distributions on Alcohol and Tobacco Use

Start by computing the mean distributions of smoking and alcohol use statesa cross
the simulation

```{r}
# Compute the proportion of each smoking status
smoking_prop <- prop.table(table(agent_dt$smoking_status))

# Compute the mean proportion of current, former, and never smokers
mean_current_smokers <- mean(smoking_prop["Current"])
mean_former_smokers <- mean(smoking_prop["Former"])
mean_never_smokers <- mean(smoking_prop["Never"])

# Print the mean smoking status distribution
cat("Mean smoking status distribution:\n")
cat("Current Smokers: ", mean_current_smokers, "\n")
cat("Former Smokers: ", mean_former_smokers, "\n")
cat("Never Smokers: ", mean_never_smokers, "\n")

# Compute the proportion of each alcohol use status
alc_prop <- prop.table(table(agent_dt$alc_use_status))

# Compute the mean proportion of each alcohol use status
mean_abstainers <- mean(alc_prop[1])
mean_occasional <- mean(alc_prop[2])
mean_regular <- mean(alc_prop[3])
mean_heavy <- mean(alc_prop[4])

# Print the mean alcohol use status distribution
cat("Mean alcohol use status distribution:\n")
cat("Abstainers: ", mean_abstainers, "\n")
cat("Occasional Users: ", mean_occasional, "\n")
cat("Regular Users: ", mean_regular, "\n")
cat("Heavy Users: ", mean_heavy, "\n")

```
Plot current smoking rate over time:

```{r}
# Define the time ticks you want to analyze
time_ticks <- c(seq(1, last_tick, by = 10), last_tick)

# Calculate the proportion of current, former, and never smokers at the specified time ticks
smokers_by_tick <- agent_dt[tick %in% time_ticks, .(
  current_smokers = sum(smoking_status == "Current") / .N,
  former_smokers = sum(smoking_status == "Former") / .N,
  never_smokers = sum(smoking_status == "Never") / .N
), by = tick]

# Reshape the data.table into a long format
smokers_long <- melt(smokers_by_tick, id.vars = "tick", variable.name = "smoking_status", value.name = "proportion")

# Update the smoking status labels
smokers_long[, smoking_status := factor(smoking_status, 
                                        levels = c("current_smokers", "former_smokers", "never_smokers"), 
                                        labels = c("Current", "Former", "Never"))]

# Plot the smoking rates for current, former, and never smokers at the specified time ticks
# Plot the smoking rates for current, former, and never smokers at the specified time ticks
ggplot(smokers_long, aes(x = tick, y = proportion, color = smoking_status, group = smoking_status)) +
  geom_line() +
  labs(title = "",
       x = "Time (Days)",
       y = "Distribution of Smoking States") +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  scale_y_continuous(limits = c(0, 0.7), breaks = seq(0, 0.7, by = 0.1))



```
Plot alcohol use states over time:

```{r}
calculate_proportions_alcohol <- function(data) {
  proportions <- prop.table(table(data$alc_use_status))
  df <- data.frame(Category = as.factor(as.character(names(proportions))),
                   Proportion = as.numeric(proportions)
  )
  df$Category <- factor(df$Category,
                        levels = c("0", "1", "2", "3"),
                        labels = c("Lifetime Abstention", "Low Risk Use", "Heavy Use", "Alcohol Use Disorder (AUD)"))
  return(df)
}


selected_ticks <- seq(1, max(agent_dt$tick), by = 10)
alcohol_summaries <- lapply(selected_ticks, function(x) {
  agents_at_tick <- agent_dt[tick == x]
  proportions <- calculate_proportions_alcohol(agents_at_tick)
  return(list("proportions" = proportions, "tick" = x))
})

alcohol_proportions_df <- do.call(rbind, lapply(alcohol_summaries, function(x) {
  df <- x$proportions
  df$Tick <- x$tick
  return(df)
}))

head(alcohol_proportions_df)

# Create the plot with the specified order of legend labels
ggplot(alcohol_proportions_df, aes(x = Tick, y = Proportion, color = Category, group = Category)) +
  geom_line() +
  labs(title = "",
       x = "Time (Days)",
       y = "Distribution of Alcohol Use States") +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  ylim(c(0,1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

ggplot(alcohol_proportions_df, aes(x = Tick, y = Proportion, color = Category, group = Category)) +
  geom_line() +
  labs(title = "",
       x = "Time (Days)",
       y = "Distribution of Alcohol Use States") +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(legend.title = element_blank())+
  ylim(c(0,1))+
  scale_y_continuous(breaks = seq(0, 1, 0.1))

```


## Joint Probability Computations of Alcohol and Tobacco Use

Joint probability of current smoking and alcohol use (i.e., not lifetime abstainer):

```{r}
# Filter the dataset to include only current smokers who are also alcohol users
current_smokers_alcohol_users_dt <- agent_dt[agent_dt$smoking_status == "Current" & agent_dt$alc_use_status %in% c(1, 2, 3),]

# Compute the proportion of individuals who are both current smokers and alcohol users
joint_prob_current_smokers_alcohol_users <- nrow(current_smokers_alcohol_users_dt) / nrow(agent_dt)

# Print the joint probability of current smokers and alcohol users
cat("Joint probability of current smokers and alcohol users: ", joint_prob_current_smokers_alcohol_users, "\n")

```

Joint probability of former or current/former smoking and alcohol use (i.e., not lifetime abstainer):

```{r}
# Filter the dataset to include only current smokers who are also alcohol users
former_or_current_smokers_alcohol_users_dt <- agent_dt[agent_dt$smoking_status %in% c("Current", "Former") & agent_dt$alc_use_status %in% c(1, 2, 3),]

# Compute the proportion of individuals who are both current smokers and alcohol users
joint_prob_former_or_current_smokers_alcohol_users <- nrow(former_or_current_smokers_alcohol_users_dt) / nrow(agent_dt)

# Print the joint probability of current smokers and alcohol users
cat("Joint probability of former or current smokers and alcohol users: ", joint_prob_former_or_current_smokers_alcohol_users, "\n")

```

Joint probability of never smoking and lifetime abstention from alcohol use:


```{r}
# Filter the dataset to include only current smokers who are also alcohol users
no_smoking_or_alcohol_users_dt <- agent_dt[agent_dt$smoking_status %in% c("Never") & agent_dt$alc_use_status %in% c(0),]

# Compute the proportion of individuals who are both current smokers and alcohol users
joint_prob_no_smoking_or_alcohol_users <- nrow(no_smoking_or_alcohol_users_dt) / nrow(agent_dt)

# Print the joint probability of current smokers and alcohol users
cat("Joint probability of never smoker and lifetime alcohol abstainer: ", 
    joint_prob_no_smoking_or_alcohol_users, "\n")

```


Joint probability of never/former smoking and lifetime abstention from alcohol use:

```{r}
# Filter the dataset to include only current smokers who are also alcohol users
former_or_never_smoking_and_no_alcohol_users_dt <- agent_dt[agent_dt$smoking_status %in% 
                                                              c("Never", "Former") &
                                                              agent_dt$alc_use_status %in% c(0),]

# Compute the proportion of individuals who are both current smokers and alcohol users
joint_prob_former_or_never_smoking_and_no_alcohol_users <- nrow(former_or_never_smoking_and_no_alcohol_users_dt) /
  nrow(agent_dt)

# Print the joint probability of current smokers and alcohol users
cat("Joint probability of former/never smoker and lifetime alcohol abstainer: ", 
    joint_prob_former_or_never_smoking_and_no_alcohol_users, "\n")

```

Joint probability of current/former smoking and lifetime abstention from
alcohol use:


```{r}
# Filter the dataset to include only current/former smokers who are lifetime abstainers from alcohol
current_former_smokers_lifetime_abstainers_dt <- agent_dt[agent_dt$smoking_status %in% c("Current", "Former") & agent_dt$alc_use_status == 0,]

# Compute the proportion of individuals who are both current/former smokers and lifetime abstainers from alcohol
joint_prob_current_former_smokers_lifetime_abstainers <- nrow(current_former_smokers_lifetime_abstainers_dt) / nrow(agent_dt)

# Print the joint probability of current/former smokers and lifetime abstainers from alcohol use
cat("Joint probability of current/former smokers and lifetime abstainers from alcohol use: ", joint_prob_current_former_smokers_lifetime_abstainers, "\n")
```


Joint probability of never smoking and using alcohol:

```{r}
# Filter the dataset to include only alcohol users who have never smoked
alcohol_users_never_smokers_dt <- agent_dt[agent_dt$alc_use_status %in% c(1, 2, 3) & agent_dt$smoking_status == "Never",]

# Compute the proportion of individuals who are alcohol users but never smokers
joint_prob_alcohol_users_never_smokers <- nrow(alcohol_users_never_smokers_dt) / nrow(agent_dt)

# Print the joint probability of alcohol users but never smokers
cat("Joint probability of alcohol users but never smokers: ", joint_prob_alcohol_users_never_smokers, "\n")

```



Compare the conditional, marginal, and joint probability distributions to the
Falk et al. (or other empirical) data.


```{r}
# Create a stacked bar chart of smoking and alcohol use status
ggplot(smoking_alc_use_melt, aes(x = smoking_status, y = value, fill = alcohol_use)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Agent Smoking and Alcohol Use Status Distribution",
       x = "Smoking Status",
       y = "Count",
       fill = "Alcohol Use Status") +
  theme_minimal() +
  scale_fill_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  guides(fill = "none")
```


```{r}
# Create a contingency table of smoking and alcohol use status
smoking_alc_use_table <- dcast(agent_dt[tick == last_tick], smoking_status ~ alc_use_status, value.var = "id", fun.aggregate = length)
smoking_alc_use_table <- smoking_alc_use_table[, .(smoking_status, `Lifetime Abstainer` = `0`, Occasional = `1`, Regular = `2`, Heavy = `3`)]
smoking_alc_use_melt <- melt(smoking_alc_use_table, id.vars = "smoking_status")

# Create a stacked bar chart of smoking and alcohol use status
ggplot(smoking_alc_use_melt, aes(x = smoking_status, y = value, fill = variable)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Agent Smoking and Alcohol Use Status Distribution",
       x = "Smoking Status",
       y = "Count") +
  scale_fill_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c"), 
                    labels = c("Lifetime Abstainer", "Occasional", "Regular", "Heavy"),
                    name = "Alcohol Use Status") +
  theme_minimal()

```
# Age Assignment at Entry

We check below the ages of new agents, 
and if their initial age assignments 
and age increments are working correctly. 

How many agents are present at the last tick:

```{r}
agent_dt[tick == last_tick, .N] 
```   
How many agents present at tick 991 entered after model initialization?

```{r}
agent_dt[tick == 991 & id > 9999, .N] 
```

The mean age of agents at time by time of entry at tick 1:
```{r}
agent_dt[tick == 1, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]


```
At tick 491:

```{r}
agent_dt[tick == 491, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]


```

And at tick 991:

```{r}
agent_dt[tick == 991, 
         .("mean_age_today"=mean(age), .N), 
                        by=.(entry_at_tick),
         ]
```

These data seem to match expectations of aging of newly entering agents.
See that at tick 1, only one agent had entered after model initialization.
and this agent was 18 years old.

At tick 491, this agent was 19 years old (491-365 years older than 18, rounded), 
and at tick 991, this agent was 21 years old (991-365 years older than 18, rounded).

So, the aging process seems to work as expected.


# Incarceration Statistics

Mean incarceration rate over time:

```{r}
selected_ticks <- c(seq(1, last_tick, by = 10), last_tick)

incarceration_summary <- agent_dt[tick %in% selected_ticks, 
                                 .(n_incarcerated = sum(current_incarceration_status == 1),
                                   n_agents = .N), 
                                 by = tick]

# Compute incarceration rate per 100,000 persons
incarceration_summary[, rate_per_100k := n_incarcerated / n_agents * 100000]

# Create the line plot
ggplot(incarceration_summary, aes(x = tick, y = rate_per_100k)) +
  geom_line() +
  theme_minimal() +
  labs(title = "",
       x = "Time (Days)",
       y = "Incarceration Rate (per 100,000 persons)")+
  ylim(0, 350)


```


Number and proportion of incarcerated persons at last tick:

```{r}
agent_dt[tick==last_tick, .N, by=c("current_incarceration_status")][,"prop":=round(N/sum(N)*100, 2)][]
```

Mean number and proportion of incarcerated persons at each tick across the simulation:


```{r}
selected_ticks <- c(seq(1, last_tick, by = 10), last_tick)


incarceration_summary <- agent_dt[tick %in% selected_ticks, 
                                 .(n_incarcerated = sum(current_incarceration_status == 1),
                                   n_agents = .N), 
                                 by = tick]
```


Break down number and proportions of incarcerated persons at last tick by race, sex and age:

```{r}
agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("race")][,"prop":=round(N/sum(N)*100, 0)][]

agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("female")][,"prop":=round(N/sum(N)*100, 0)][]

agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("age_groups")][,"prop":=round(N/sum(N)*100, 0)][]

agent_dt[tick==last_tick & current_incarceration_status == 1, 
         .N, 
         by=c("race", "female", "age_groups")][order(race, female, age_groups)][,"prop":=round(N/sum(N)*100, 0)][]
```

How many persons have been ever incarcerated?

To answer this question, we can group the last incarceration times by agent ID.
If last incarceration time is -1, the person has never been incarcerated. 
If last incarceration time is > -1, the person has been incarcerated at least once. 

```{r}
# Group data by agent ID and compute the last incarceration time
last_incarceration_time <- agent_dt[, 
                                    .(last_incarceration_time = max(last_incarceration_tick)), 
                                    by = id]

# Count the number of agents who have been incarcerated at least once
n_ever_incarcerated <- sum(last_incarceration_time$last_incarceration_time != -1)

# Compute the proportion of agents who have been incarcerated at least once
prop_ever_incarcerated <- n_ever_incarcerated / nrow(last_incarceration_time)


# Agents who have been incarcerated at least once
ever_incarcerated_times <- last_incarceration_time[last_incarceration_time > 1]
dim(ever_incarcerated_times)

# Extract demographic information for ever incarcerated agents
ever_incarcerated_ids <-
  unique(agent_dt[id %in% ever_incarcerated_times$id, id])
ever_incarcerated_info <- agent_dt[id %in% ever_incarcerated_ids,
                                   .SD[.N, list(
                                     id,
                                     age = last(age),
                                     race = last(race),
                                     female = last(female)
                                   )],
                                   by = id]

# Compute distributions
## mean age
age_summary <- ever_incarcerated_info[, .(mean_age=mean(age),
                                       min_age=min(age),
                                       max_age=max(age)
                                       )
                                   ]
sex_dist <-
  ever_incarcerated_info[, .N, by = female][,  .(
    Sex = c("Female", "Male"),
    Count = c(N[female == 0], N[female == 1]),
    "Proportion of ever incarcerated" = round(N / sum(N), 2)
  )]
print(sex_dist)

race_dist <- ever_incarcerated_info[, .N, by = race][,
  .(Race = race, 
    Count = N,
    Proportion = round(N / sum(N), 2)
   )
  ]
race_dist[]

```

Thus, `r n_ever_incarcerated` persons have been incarcerated at least once. The
proportion of persons who have been ever incarcerated is `r prop_ever_incarcerated`.

Now let us compute the number of females and males whoa have ever been incarcerated,
and the proportions by race (% of White, Black, ...)

```{r}
first_tick_race_sex <-
  agent_dt[, .SD[1], by = id][, .(
    ID = id,
    Tick = tick,
    Female = female,
    Race = race
  )]

# Count number of females and males
sex_count <- first_tick_race_sex[, .N, by = Female][,
                                               .(Sex = c("Female", "Male"), 
                                                 Count = c(N[Female == 1], N[Female == 0]))                                               ]
print(sex_count)

# Count number of individuals in each race
race_count <- first_tick_race_sex[, .N, by = Race][,
                                                   .(Race, Count = N)]
print(race_count)
```

Thus, the distribution of ever incarcerated persons by sex is

```{r}
cbind(sex_count$Sex, round(as.numeric(sex_dist$Count/sex_count$Count), 3))
```

The distribution of ever incarcerated persons by race is:

```{r}
cbind(race_count$Race, round(as.numeric(race_dist$Count/race_count$Count), 3))
```

What is the duration of incarceration for persons who have been ever incarcerated? 
Restricting our data to only persons who have been released from their last
incarceration episode, we can estimate this as:

```{r}
# Filter the data to include only the observations for agents who have been ever incarcerated and released
ever_incarcerated_released <-
  agent_dt[last_release_tick > last_incarceration_tick,
           .SD[1, ], # ensure times for each incarceration, release episode appear only once
           by = .(id, last_incarceration_tick)]

# Compute the length of incarceration time for each ever incarcerated and released agent
ever_incarcerated_released[, incarceration_length := last_release_tick - last_incarceration_tick,
                           by = id]

# Compute the summary statistics of the lengths of incarceration time for the ever incarcerated and released agents
incarceration_summary <-
  ever_incarcerated_released[, .(
    min_incarceration_length = min(incarceration_length),
    median_incarceration_length = median(incarceration_length),
    max_incarceration_length = max(incarceration_length)
  ),
  by = id]

# Compute the distribution of incarceration times for all incarceration episodes followed by release
summary(ever_incarcerated_released$incarceration_length)
```

Assess the total number of incarceration-release events and the number of unique
agents who experienced them:

```{r}
length(ever_incarcerated_released$id)
length(unique(ever_incarcerated_released$id))
```

Distribution of incarceration events:
```{r}
hist(ever_incarcerated_released$incarceration_length, 
     breaks = seq(0, 3000, by = 30),
     xlab = "Length of incarceration (days)",
     ylab = "Number of individuals",
     main = "Distribution of incarceration lengths")
```

Break down number and proportions of persons ever incarcerated at last tick by race, sex and age:

```{r}
ever_inc_dt <- 
  agent_dt[tick==last_incarceration_tick & n_incarcerations > 0, 
           .N, 
           by=c("race", "female", "age_groups")][order(race, female, age_groups)][
             ,"prop":=round(N/sum(N)*100, 0)][]

ever_inc_dt
```
```{r}
ever_inc_dt[, colSums(.SD), .SDcols = 4:5]
```

Mean number of incarcerations at last tick:

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_incarcerations" = mean(n_incarcerations),
           "min_n_incarcerations" = min(n_incarcerations),
           "max_n_incarcerations" = max(n_incarcerations),
           "median_n_incarcerations" = median(n_incarcerations),
           "n_ever_incarcerated_agents"=sum(n_incarcerations > 0),
         .N)] 
```   

## Incarceration Duration

Compute statistics on agents whose last release date is 
after their last incarceration date.

```{r}
agent_dt[tick == last_tick & last_release_tick >= last_incarceration_tick, 
         .(
           "mean_duration" = mean(last_release_tick - last_incarceration_tick),
           "max_duration" = max(last_release_tick - last_incarceration_tick),
           "min_duration" = min(last_release_tick - last_incarceration_tick),
           "median_duration" = median(last_release_tick - last_incarceration_tick),
         .N)] 
```



# Release Statistics

## Number

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_releases" = mean(n_releases),
           "min_n_releases" = min(n_releases),
           "max_n_releases" = max(n_releases),
           "median_n_releases" = median(n_releases),
           "n_ever_released_agents"=sum(n_releases > 0),
         .N)] 
```
## Duration
Compute statistics on agents whose last release date is 
\*before\* their last incarceration date.

```{r}
agent_dt[tick == last_tick & last_release_tick < last_incarceration_tick, 
         .(
           "mean_rel_dur" = mean(abs(last_release_tick - last_incarceration_tick)),
           "max_rel_dur" = max(abs(last_release_tick - last_incarceration_tick)),
           "min_rel_dur" = min(abs(last_release_tick - last_incarceration_tick)),
           "median_rel_dur" = median(abs(last_release_tick - last_incarceration_tick)),
         .N)] 
```


# Smoking transition behavior

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_smkg_trans" = mean(n_smkg_stat_trans),
           "max_n_smkg_trans" = max(n_smkg_stat_trans),
           "min_n_smkg_trans" = min(n_smkg_stat_trans),
           "median_n_smkg_trans" = median(n_smkg_stat_trans),
         .N)] 
```

Smoking State Distributions:

```{r}
ans <- agent_dt[tick == last_tick, .N, by=c("race", "smoking_status")][order(race)]
ans 

ans[, prop := N / sum(N), by = race][]

ans_num_smokers <- ans[smoking_status %in% c("Former", "Current", "Never"),
                       .(num_smokers = sum(N)),
                       by = race]
ans_num_smokers
sum(ans_num_smokers$num_smokers)

sim_prop_smokers_by_race <- ans_num_smokers$num_smokers/sum(ans_num_smokers$num_smokers)


```


# Alcohol transition behavior

```{r}
agent_dt[tick == last_tick, 
         .(
           "mean_n_alc_trans" = mean(n_alc_use_stat_trans),
           "max_n_alc_trans" = max(n_alc_use_stat_trans),
           "min_n_alc_trans" = min(n_alc_use_stat_trans),
           "median_n_alc_trans" = median(n_alc_use_stat_trans),
         .N)] 
```

Alcohol Use State Distributions:

```{r}
ans <- agent_dt[tick == last_tick, .N, by=alc_use_status]
prop <- ans$N/sum(ans$N)
cbind(ans[,1], prop)
```

# Smoking and Alcohol Use among Incarcerated Persons vs Not-

## Smoking

Compare the proportion of the current smokers (to all persons) at the last tick 
in the n_releases = 0 group vs the n_releases = 1 group.

```{r}
agent_dt[tick == last_tick, .N, by = n_releases][]
agent_dt[tick == last_tick, .N, by = smoking_status][]
agent_dt[tick == last_tick, .N, by = c("smoking_status", "n_releases")][]


# Filter data at the last_tick
last_tick_data <- agent_dt[tick == last_tick]

# Calculate the ratio for each group
ratios <- last_tick_data[, .(
  n_current_smokers = sum(smoking_status == "Current"),
  n_total_persons = sum(smoking_status %in% c("Current", "Former", "Never"))
), by = .(group = n_releases >= 1)]

# Compute the ratio of current smokers to total smokers in each group
ratios[, ratio := n_current_smokers / n_total_persons]

# Print the results
ratios



```


The population size at each time tick is 

```{r}
sum(agent_dt[tick == last_tick, .N, by = c("smoking_status", "n_releases")][["N"]])
```

Now let us compute this comparison-of-ratios across all time steps, i.e.,
mean across all time steps of the current smoking proportion in the 
n_releases = 0 group vs. the n_releases >= 1 group. We add
the standard deviation and the interquartile range as measures of variation
across both groups.

```{r}
<<<<<<< HEAD
# Calculate the ratio for each group and each tick
ratios_by_tick <- agent_dt[, .(
  n_current_smokers = sum(smoking_status == "Current"),
  n_total_persons = sum(smoking_status %in% c("Current", "Former", "Never"))
), by = .(tick, group = n_releases >= 1)]

# Compute the ratio of current smokers to total smokers in each group and each tick
ratios_by_tick[, ratio := n_current_smokers / n_total_persons]

# Calculate the mean ratio, standard deviation, and interquartile range across all time steps for each group
summary_by_group <- ratios_by_tick[, .(
  mean_ratio = mean(ratio, na.rm = TRUE),
  sd_ratio = sd(ratio, na.rm = TRUE),
  iqr_ratio = IQR(ratio, na.rm = TRUE)
), by = group]

# Print the results
summary_by_group
```

Test if the means for the two groups are statistically different:

```{r}
# Extract the ratio data for each group
group_0_ratios <- ratios_by_tick[group == FALSE, ratio]
group_1_ratios <- ratios_by_tick[group == TRUE, ratio]
=======
n_releases_0 <- current_smoking_by_release_dcast[,2]
#n_releases_ge1 <- current_smoking_by_release_dcast[,3:6]

n_releases_0[1]/sum(n_releases_0[1:2])
#sum(n_releases_ge1[1,])/sum(n_releases_ge1[1:2,])
>>>>>>> network-reparameterization

# Perform the two-sample t-test
t_test_result <- t.test(group_0_ratios, group_1_ratios)

# Print the t-test result
t_test_result

```

### Compute change in current smoking rate over time since release


```{r}
calculate_proportions_smoking <- function(data) {
  proportions <- prop.table(table(data$smoking_status))
  df <- data.frame(Category = names(proportions),
                   Proportion = as.numeric(proportions))
  df$Category <- factor(df$Category,
                        levels = c("Never", "Former", "Current"),
                        labels = c("Never", "Former", "Current"))
  return(df)
}

summary_after_release <- function(time_period) {
  agents_within_time_period <- agent_dt[last_release_tick > 0 &
                                          tick - last_release_tick <= time_period]

  proportions <- calculate_proportions_smoking(agents_within_time_period)
  agent_count <- nrow(agents_within_time_period)
  return(list("proportions" = proportions, "agent_count" = agent_count))
}

time_periods <- c(1, 7, 14, 30, 90, 180, 365)
labels <- c("1D", "1W", "2W", "1M", "3M", "6M", "1Y")
summaries_smoking <- lapply(time_periods, function(x) summary_after_release(x))


summaries_smoking



```
Plot the rates of agents who are current smokers:

```{r}
class(summaries_smoking)

current_smoker_proportions <- sapply(
  summaries_smoking, 
  function(x) x$proportions[x$proportions$Category == "Current", "Proportion"])

current_smoker_df <- data.frame(Time = time_periods,
                                Labels = labels,
                                Proportion = current_smoker_proportions)

ggplot(current_smoker_df, aes(x = Time, y = Proportion, group = 1)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = time_periods, labels = labels) +
  labs(title = "",
       x = "Time After Release",
       y = "Proportion of Current Smokers") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Alcohol Use

Compare the alcohol use rates in the: (1) entire agent population; 
(2) agent population that has been never released;
(3) agent population has been released at least once:

```{r}
# Entire agent population
all_agents <- agent_dt

# Agents who have never been released (n_releases == 0)
never_released_agents <- agent_dt[n_releases == 0]

# Agents who have been released at least once (n_releases >= 1)
released_agents <- agent_dt[n_releases >= 1]


# Function to calculate proportions
calculate_proportions <- function(data) {
  proportions <- prop.table(table(data$alc_use_status))
  df <- data.frame(Category = names(proportions),
                   Proportion = as.numeric(proportions))
  df$Category <- factor(df$Category,
                        levels = c(0, 1, 2, 3),
                        labels = c("Lifetime Abstainer", "Occasional User", "Regular User", "Heavy User"))
  return(df)
}

# Entire agent population
all_agents_summary <- calculate_proportions(all_agents)

# Agents who have never been released
never_released_agents_summary <- calculate_proportions(never_released_agents)

# Agents who have been released at least once
released_agents_summary <- calculate_proportions(released_agents)

# Display summaries
print("Summary for Entire Agent Population")
print(all_agents_summary)

print("Summary for Agents Who Have Never Been Released")
print(never_released_agents_summary)

print("Summary for Agents Who Have Been Released at Least Once")
print(released_agents_summary)
```
Once agents are released, the baseline state transition probabilities take effect,
which force the alcohol use rates to equilibrate at the baseline state distribution of alcohol use. 
This might be the reason why the heavy alcohol use rates do not differ between the groups
that have never been released vs. the group that has been released at least once. 


To account for this effect of baseline transition probabilities, 
we will instead compare the alcohol use rates shortly after release,
i.e., we will compute the alcohol use rates 
for a certain period after the agents are released. 
This may help uncover the immediate effect of release on alcohol use rates.

```{r}
# Define a one-year period after release
fixed_time_after_release <- 1

# Filter agents who were released at least once and within one year after release
within_fixed_time_after_release_agents <- agent_dt[last_release_tick > 0 & tick <= last_release_tick + fixed_time_after_release]

# Calculate proportions for agents within one year after release
within_fixed_time_after_release_summary <- calculate_proportions(within_fixed_time_after_release_agents)

# Display the summary
print("Summary for Agents Within One Year After Release")
print(within_fixed_time_after_release_summary)

# Note for 
```
The proportion of heavy users one day after release is in the right range:
approximately `r within_fixed_time_after_release_summary[within_fixed_time_after_release_summary$Category == "Heavy User", "Proportion"] * 100`%.
But the proportion of lifetime abstainers at `r within_fixed_time_after_release_summary[within_fixed_time_after_release_summary$Category == "Lifetime Abstainer", "Proportion"] * 100`
seems higher than it should be.

Sensitivity Analysis of time since release:

```{r}
# Function to filter agents and calculate proportions for the given time period after release
summary_after_release <- function(time_period) {
  agents_within_time_period <- agent_dt[last_release_tick > 0 & tick <= last_release_tick + time_period]
  proportions <- calculate_proportions(agents_within_time_period)
  return(proportions)
}

# Time periods in ticks (days): 1 day, 1 week, 2 weeks, 1 month, 3 months, 6 months, 1 year
time_periods <- c(1, 7, 14, 30, 90, 180, 365)

# Calculate summaries for each time period
summaries <- lapply(time_periods, function(x) summary_after_release(x))

# Display summaries
names(summaries) <- paste(time_periods, "days after release")
summaries

```


```{r}
time_points <- c(1, 7, 14, 30, 90, 180, 365)
labels <- c("1D", "1W", "2W", "1M", "3M", "6M", "1Y")

heavy_user_proportions <- sapply(summaries, function(x) x[x$Category == "Heavy User", "Proportion"])

heavy_user_df <- data.frame(Time = time_points,
                            Labels = labels,
                            Proportion = heavy_user_proportions)

ggplot(heavy_user_df, aes(x = Time, y = Proportion, group = 1)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = time_points, labels = labels) +
  labs(title = "",
       x = "Time After Release",
       y = "Proportion of Agents with AUD") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

