---
title: "Incarceration-Alcohol Use interaction"
output: html_notebook
bibliography: cadre-appendix-library.bib
---

```{r, results='hide', echo=FALSE, message=FALSE}
# Libraries ----------
rm(list=ls())
library(ergm)
library(network)
library(dplyr)

```

```{r, results='hide', echo=FALSE, message=FALSE}
# Libraries ----------
load("intialized-network.RData")
```

We start with an initialized population with the following attributes:

```{r, results='hide', echo=FALSE, message=FALSE}
list.vertex.attributes(net)
```

Assign alcohol use state distributions to the population:

```{r}
abstainer <- 8.3/100  #state 1
occasional_drinker <- 72.9/100 #state 2
regular_drinker <- 13.2/100 #state 3
high_risk_or_AUD <- 1-(abstainer+occasional_drinker+regular_drinker) #state 4

alcohol.use.states <- c(1, 2, 3, 4)

alcohol.use.behavior <- sample(alcohol.use.states, n, 
                               replace=TRUE, 
                               prob = c(abstainer, occasional_drinker,
                                        regular_drinker,high_risk_or_AUD)
                               )
table(alcohol.use.behavior)
table(alcohol.use.behavior)/n

net %v% "alcohol.use.behavior" <- alcohol.use.behavior
table(net %v% "alcohol.use.behavior", exclude = NULL)
```

To simulate one-step transitions between the states, define transition probabilities to "level up" or down. 

```{r, results='hide', echo=FALSE, message=FALSE}
# level up
trans.prob.0.1 <- 0/100 
trans.prob.1.2 <- 10/100
trans.prob.2.3 <- 10/100

# level down
trans.prob.1.0 <- 5/100
trans.prob.2.1 <- 5/100
trans.prob.3.2 <- 5/100


alc.state.transition.matrix <- 
  matrix(
    c(
      1, trans.prob.0.1, 0, 0,
      trans.prob.1.0, 1-(trans.prob.1.0+trans.prob.1.2), trans.prob.1.2, 0,
      0, trans.prob.2.1, 1-(trans.prob.2.1+trans.prob.3.2), trans.prob.2.3,
      0, 0, trans.prob.3.2, 1-trans.prob.3.2),
      nrow=4, byrow = TRUE
    )
  

alc.state.transition.matrix
```

Write a function to simulate alcohol use transitions. 
Agent cannot both level up and down at one time step. Half the agents are eligible to step up, 
and half are eligible to step down. A decision is made to
either level up (or stay in the same state), or level down (or stay in the same state). 

```{r}
simulate.alcohol.use.transitions <- function (net, verbose = FALSE){
n <- network.size(net)
    for (agent in 1:n){
    # record agent alcohol use state
    alcohol.use.behavior.agent <- (net %v% "alcohol.use.behavior")[agent] 
    
    if (level.up.or.down[agent] < 0.5 &&  alcohol.use.behavior.agent  != 1){
      if (verbose) {cat("Transition down??? ")}
      if (level.up.or.down[agent] <= alc.state.transition.matrix[alcohol.use.behavior.agent,                                                            alcohol.use.behavior.agent-1]){
        set.vertex.attribute(net, "alcohol.use.behavior", alcohol.use.behavior.agent-1, agent)
         if (verbose) {cat(" Transitioning down!!!", "\n")}
      } else {
         if (verbose) {cat("Not transitioning down!!!", "\n")}
      }
    } else if (level.up.or.down[agent] > 0.5 &&  alcohol.use.behavior.agent != 4){
       if (verbose) {cat("Transition up??? ")}
      if (level.up.or.down[agent] <= alc.state.transition.matrix[alcohol.use.behavior.agent, alcohol.use.behavior.agent+1]){
        set.vertex.attribute(net, "alcohol.use.behavior", alcohol.use.behavior.agent+1, agent)
         if (verbose) {cat(" Transitioning up!!!", "\n")}
      } else {
         if (verbose) {cat("Not transitioning up!!!", "\n")}
      }
    }
  }
return (net)
  }
```
Simulate the above function for many time steps:

```{r}
max.time <- 10
n <- network.size(net)
 
for (time in 1:10){
  cat("Entering time step ", time, "\n")
  level.up.or.down <- runif(n, 0, 1)
  net.post.sim <- simulate.alcohol.use.transitions(net)
}

summary(net %v% "alcohol.use.behavior")
summary(net.post.sim %v% "alcohol.use.behavior")

summary((net.post.sim %v% "alcohol.use.behavior") - net %v% "alcohol.use.behavior")
```
